---
title: "Genetic_load_figures_Ppinaster_study"
author: "Thomas Francisco"
date: "2026-04-01"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true 
    collapsed: True
    toc_depth: 3
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/tfrancisco/Documents/Thesis/")
rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
```

```{r library, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(writexl)
library(ggplot2)
library(tidyverse)
```

# Introduction

This script analyses genetic load results in *P. pinaster* adults and juveniles in the Lacanau and Tocchi populations. The aim was to investigate the levels of number of deleterious mutations in homozygosis and heterozygosis across populations and to gain insight into the dynamics of these loads across cohorts. 

The results were obtained from the genetic load calculation script using SNPeff and Provean softwares.

We computed each load type: realised, potential and total loads with and without correction for population genetic diversity.

# Loading data

```{r genetic load data}
Df_load_total_realised <- read.csv("Data/Chapter3/Pinus_pinaster/genetic_load/Ppinaster_chapter3_GL_all_stats_new.csv",h=T,sep=";",dec=",")

Df_load_total_realised_num <- data.frame(Indiv_ID=Df_load_total_realised$Indiv_ID,dplyr::mutate_all(Df_load_total_realised[,-1], function(x) as.numeric(as.character(x))))

#Load info on adults and juveniles
load("Results/Chapter3/Pinaster/Comparison_juveniles_adults/Data/combined_rda_data_Lacanau_LC.Rdata")
combined_rda_data_Lacanau_LC$Indiv_ID <- row.names(combined_rda_data_Lacanau_LC)

#Change name to match loads dataframe
combined_rda_data_Lacanau_LC$Indiv_ID <- ifelse(
  grepl("_0*(\\d{1,3})$", combined_rda_data_Lacanau_LC$Indiv_ID) & as.numeric(sub(".*_(\\d+)$", "\\1", combined_rda_data_Lacanau_LC$Indiv_ID)) <= 500,
  paste0(
    "FR_PP_62_", 
    sub(".*_(\\d+)$", "\\1", combined_rda_data_Lacanau_LC$Indiv_ID), 
    "-", 
    combined_rda_data_Lacanau_LC$Indiv_ID
  ),
  combined_rda_data_Lacanau_LC$Indiv_ID
)

load("Results/Chapter3/Pinaster/Comparison_juveniles_adults/Data/combined_rda_data_Tocchi_LC.Rdata")
combined_rda_data_Tocchi_LC$Indiv_ID <- row.names(combined_rda_data_Tocchi_LC)

#Change name to match loads dataframe
num_suffix <- as.numeric(sub(".*_(\\d+)$", "\\1", combined_rda_data_Tocchi_LC$Indiv_ID))

#Only transform 1 to 500 to match label used in the other dataframe
combined_rda_data_Tocchi_LC$Indiv_ID <- ifelse(
  num_suffix <= 500,
  paste0(
    "IT_PP_63_", 
    sprintf("%02d", num_suffix),
    "-", 
    combined_rda_data_Tocchi_LC$Indiv_ID
  ),
  combined_rda_data_Tocchi_LC$Indiv_ID
)
```

Add the cohort information
```{r age of trees}
Df_load_total_realised_Lacanau <- merge(Df_load_total_realised_num,combined_rda_data_Lacanau_LC,"Indiv_ID")

Df_load_total_realised_Tocchi <- merge(Df_load_total_realised_num,combined_rda_data_Tocchi_LC,"Indiv_ID")
```

```{r save for supplementary file}
Df_load_total_realised_Tocchi_correct_name <- 
Df_load_total_realised_Tocchi_order <- 

  write.csv(Df_load_total_realised_Lacanau, file="Results/Chapter3/Pinaster/Genetic_load_new_estimates/numbers/Df_load_total_realised_Lacanau.csv")
```

# Realised load

The realised genetic load is expected to currently have an impact on the individual fitness as it corresponds to the expected deleterious mutations in homozygosis. 

## Ratio of realised load by dividing with the number of neutral mutations in homozygosis

```{r Boxplot}
list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_realised_load <- get(paste0("Df_load_total_realised_",plot))
  
  t_test_result <- t.test(Recessive_GLr ~ Group, data = df_realised_load)

    assign(paste0("realised_load_difference_across_adults_juv_",plot),t_test_result)
  
  #Plot
realised_load_boxplot <- ggplot(df_realised_load, aes(x = Group, y = Recessive_GLr, fill = Group)) +
  geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +
  scale_color_manual(values = c("darkred", "darkgreen")) +
  labs(
    title = paste0("Boxplot Plot of realised Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "realised Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.12, 0.18)) + 
  scale_y_continuous(breaks = seq(0.12, 0.18, by = 0.002)) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(realised_load_boxplot)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/realised/figures/GL_realised_boxplot_",plot,".pdf"), width = 8, height = 6)
print(realised_load_boxplot)
dev.off()

realised_load_boxplot_wo_points <- ggplot(df_realised_load, aes(x = Group, y = Recessive_GLr, fill = Group)) +
  #geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) + #points
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +
  scale_fill_manual(values = c("darkred", "darkgreen")) +
  scale_color_manual(values = c("darkred", "darkgreen")) + 
  labs(
    title = paste0("Boxplot Plot of realised Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "realised Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.12, 0.18)) +   
  scale_y_continuous(breaks = seq(0.12, 0.18, by = 0.02)) + 
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(realised_load_boxplot_wo_points)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/realised/figures/GL_realised_boxplot_wopoints_",plot,".pdf"), width = 8, height = 6)
print(realised_load_boxplot_wo_points)
dev.off()

}

print(realised_load_difference_across_adults_juv_Lacanau)
print(realised_load_difference_across_adults_juv_Tocchi)
```

## Raw number of deleterious mutations in heterozygosis

We computed the realised load without normalising by the number of neutral homozygous sites because comparisons were performed among individuals within the same population, which are expected to share a common demographic history and similar levels of neutral genetic variation.

Graph with final colors.
```{r graph with good colors}
list_plots <- c("Lacanau","Tocchi")

#Define color lists
color_list <- list(
  Lacanau = c(
    Adult = "#3A6FB0",
    Juvenile = "#A7C7E7"
  ),
  Tocchi = c(
    Adult = "#C49A00",
    Juvenile = "#FFD700"
  )
)
 
for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_realised_load <- get(paste0("Df_load_total_realised_",plot))
  
  #Run t-test
  t_test_result <- t.test(Homozygous_deleterious_counts ~ Group, data = df_realised_load)
  assign(paste0("raw_realised_load_difference_across_adults_juv_",plot), t_test_result)
  
  #Select correct color palette
  col_pal <- color_list[[plot]]
  
  #Boxplot
  raw_realised_load_boxplot <- ggplot(df_realised_load, 
                                       aes(x = Group, 
                                           y = Homozygous_deleterious_counts, 
                                           fill = Group)) +
    geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = col_pal) +
    scale_color_manual(values = col_pal) +
    labs(
      title = paste0("Boxplot of realised Raw Genetic Load (Provean) - ", plot),
      y = "Number of deleterious homozygous mutations"
    ) +
    coord_cartesian(ylim = c(60, 125)) +
    scale_y_continuous(breaks = seq(60, 125, by = 10)) +
       theme_bw() +
theme(
  panel.grid.minor.y = element_blank(),  
  panel.grid.minor.x = element_blank(),   
  panel.grid.major.y = element_line(color="grey90"),
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)

  plot(raw_realised_load_boxplot)

  #Save plot
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/realised/figures/GL_raw_realised_boxplot_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_realised_load_boxplot)
  dev.off()
  
  #Boxplot without individual points
  raw_realised_load_boxplot_wo_points <- ggplot(df_realised_load, 
                                                 aes(x = Group, 
                                                     y = Homozygous_deleterious_counts, 
                                                     fill = Group)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = col_pal) +
    labs(
      title = paste0("Boxplot of realised Raw Genetic Load (Provean) - ", plot),
      y = "Number of deleterious homozygous mutations"
    ) +
    coord_cartesian(ylim = c(60, 125)) +
    scale_y_continuous(breaks = seq(60, 125, by = 10)) +
    theme_bw() +
      theme_bw() +
theme(
  panel.grid.minor.y = element_blank(),  
  panel.grid.minor.x = element_blank(),   
  panel.grid.major.y = element_line(color="grey90"),  
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)

  plot(raw_realised_load_boxplot_wo_points)
  
  #Save plot without points
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/realised/figures/GL_raw_realised_boxplot_wopoints_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_realised_load_boxplot_wo_points)
  dev.off()
}

#Print t-test results
print(raw_realised_load_difference_across_adults_juv_Lacanau)
print(raw_realised_load_difference_across_adults_juv_Tocchi)
```

Combined two populations into one boxplot
```{r plot full legend}
#Merge the two datasets and create a new combined category
df_all <- rbind(
  transform(Df_load_total_realised_Lacanau, population = "Lacanau"),
  transform(Df_load_total_realised_Tocchi, population = "Tocchi")
)

#Create combined factor for x-axis: Population + Age
df_all$pop_age <- factor(
  paste(df_all$population, df_all$Group, sep = "_"),
  levels = c("Lacanau_Adult","Lacanau_Juvenile",
             "Tocchi_Adult","Tocchi_Juvenile")
)

#Colors
custom_colors <- c(
  "Lacanau_Adult"     = "#3A6FB0",   
   "Lacanau_Juvenile" = "#A7C7E7",   
  "Tocchi_Adult"      = "#E6C300",    
    "Tocchi_Juvenile"  = "#FFF4A3"   
)

legend_colors <- c(
  "Lacanau" = "#3A6FB0",
  "Tocchi"  = "#E6C300"
)

raw_realised_load_boxplot_wo_points_combined <- ggplot(df_all, aes(x = pop_age, y = Homozygous_deleterious_counts, fill = pop_age)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = custom_colors) +
    labs(
      title = "Boxplot of realised Raw Genetic Load (Provean) - ",
      y = "Number of deleterious homozygous mutations"
    ) +
    coord_cartesian(ylim = c(50, 125)) +
    scale_y_continuous(breaks = seq(50, 125, by = 10)) +
    theme_bw() +
      theme_bw() +
theme(
  panel.grid.minor.y = element_blank(),  
  panel.grid.minor.x = element_blank(),   
  panel.grid.major.y = element_line(color="grey90"),  
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)

  plot(raw_realised_load_boxplot_wo_points_combined)
  
  #Save plot without points
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/realised/figures/GL_raw_realised_boxplot_wopoints_both_plots_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_realised_load_boxplot_wo_points_combined)
  dev.off()
```

```{r significant differences realised load raw}
#one-way anova to see if groups are differents
anova_result_realised_raw <- aov(Homozygous_deleterious_counts ~ pop_age, data = df_all)
summary(anova_result_realised_raw)

#They are different so we can do a Tukey post hoc to see which groups are different
tukey_result <- TukeyHSD(anova_result_realised_raw)
tukey_result

# -> They are all differents

#Let's see if Lacanau population (adults+ offsprings) is different than the Tocchi population (adults + offsprings)
anova_result_pop_realised_raw <- aov(Homozygous_deleterious_counts ~ population, data = df_all)
summary(anova_result_pop_realised_raw)

#compute median per group
df_all %>% group_by(population) %>% summarise(median = median(Homozygous_deleterious_counts))
```

# Total load

The total load correspond to the ratio between the number of deterious mutatations in homozygosis and heterozygosis/ by the number of neutral mutations in both homo- and heterozygosis or to the raw number without dividing by neutral mutations.

## Ratio of total load by dividing with the total number of neutral mutations

```{r Boxplot total load}
list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_total_load <- get(paste0("Df_load_total_realised_",plot))
  
  t_test_result <- t.test(Total_.allele_counts._GLt ~ Group, data = df_total_load)

    assign(paste0("total_load_difference_across_adults_juv_",plot),t_test_result)
 
  #Plot
total_load_boxplot <- ggplot(df_total_load, aes(x = Group, y = Total_.allele_counts._GLt, fill = Group)) +
  geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +
  scale_color_manual(values = c("darkred", "darkgreen")) +
  labs(
    title = paste0("Boxplot Plot of Total Genetic Load estimates from Provean ", plot),
   #x = "Group-age",
    y = "Total Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.12, 0.18)) +  
  scale_y_continuous(breaks = seq(0.12, 0.18, by = 0.02)) +  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(total_load_boxplot)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Total/figures/GL_total_boxplot_",plot,".pdf"), width = 8, height = 6)
print(total_load_boxplot)
dev.off()

total_load_boxplot_wo_points <- ggplot(df_total_load, aes(x = Group, y = Total_.allele_counts._GLt, fill = Group)) +
  #geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) + #points
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +  
  scale_fill_manual(values = c("darkred", "darkgreen")) +
  scale_color_manual(values = c("darkred", "darkgreen")) +  
  labs(
    title = paste0("Boxplot Plot of Total Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "Total Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.12, 0.18)) + 
  scale_y_continuous(breaks = seq(0.12, 0.18, by = 0.02)) + 
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(total_load_boxplot_wo_points)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Total/figures/GL_total_boxplot_wopoints_",plot,".pdf"), width = 8, height = 6)
print(total_load_boxplot_wo_points)
dev.off()
}

print(total_load_difference_across_adults_juv_Lacanau)
print(total_load_difference_across_adults_juv_Tocchi)
```

## Number of total deleterious mutations

```{r Boxplot raw total load}
list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_total_load <- get(paste0("Df_load_total_realised_",plot))
  
  t_test_result <- t.test(Allele_deleterious_counts ~ Group, data = df_total_load)

    assign(paste0("raw_total_load_difference_across_adults_juv_",plot),t_test_result)
 
  #Plot
raw_total_load_boxplot <- ggplot(df_total_load, aes(x = Group, y = Allele_deleterious_counts, fill = Group)) +
  geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +
  scale_color_manual(values = c("darkred", "darkgreen")) + 
  labs(
    title = paste0("Boxplot Plot of Raw Total Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "Number of deleterious mutations (Provean)"
  ) +
  coord_cartesian(ylim = c(240, 320)) +  
  scale_y_continuous(breaks = seq(240, 320, by = 20)) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(raw_total_load_boxplot)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Total/figures/GL_raw_total_boxplot_",plot,".pdf"), width = 8, height = 6)
print(raw_total_load_boxplot)
dev.off()

raw_total_load_boxplot_wo_points <- ggplot(df_total_load, aes(x = Group, y = Allele_deleterious_counts, fill = Group)) +
  #geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) + #points
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) + 
  scale_color_manual(values = c("darkred", "darkgreen")) + 
  labs(
    title = paste0("Boxplot Plot of Raw Total Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "Number of deleterious mutations (Provean)"
  ) +
  coord_cartesian(ylim = c(240, 320)) +  
  scale_y_continuous(breaks = seq(240, 320, by = 20)) +  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(raw_total_load_boxplot_wo_points)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Total/figures/GL_raw_total_boxplot_wopoints_",plot,".pdf"), width = 8, height = 6)
print(raw_total_load_boxplot_wo_points)
dev.off()
}

print(raw_total_load_difference_across_adults_juv_Lacanau)
print(raw_total_load_difference_across_adults_juv_Tocchi)
```

# Potential load

## Ratio of potential load by dividing with the number of neutral heterozygous mutations

Potential load correspond to the number of deleterious mutations in heterozygous to the number of neutral mutations in heterozygous.

GLpot = (total number of deleterious mutations – 2 × number of homozygous deleterious mutations)/(total number of neutral mutations – 2 × number of homozygous neutral mutations).

```{r compute potential load}
list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_potential_load <- get(paste0("Df_load_total_realised_",plot))
  
df_potential_load$Number_hetero_mut_GLp <- df_potential_load$Allele_neutral_counts - (2*df_potential_load$Homozygous_.neutral_counts)

df_potential_load$Number_hetero_deleterious_mut_GLp <- df_potential_load$Allele_deleterious_counts - (2*df_potential_load$Homozygous_deleterious_counts)

df_potential_load$Provean_GLp_Ppinaster_Set2 <- df_potential_load$Number_hetero_deleterious_mut_GLp / (df_potential_load$Number_hetero_mut_GLp)

assign(paste0("Df_load_total_realised_",plot),df_potential_load)
}
```

```{r Boxplot potential load}
list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_potential_load <- get(paste0("Df_load_total_realised_",plot))
  
  t_test_result <- t.test(Provean_GLp_Ppinaster_Set2 ~ Group, data = df_potential_load)

    assign(paste0("potential_load_difference_across_adults_juv_",plot),t_test_result)
 
  #plot
potential_load_boxplot <- ggplot(df_potential_load, aes(x = Group, y = Masked_GLm, fill = Group)) +
  geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) +
  scale_fill_manual(values = c("darkred", "darkgreen")) +
  scale_color_manual(values = c("darkred", "darkgreen")) +  
  labs(
    title = paste0("Boxplot Plot of Potential Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "Potential Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.10, 0.22)) +  
  scale_y_continuous(breaks = seq(0.10, 0.22, by = 0.02)) +  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(potential_load_boxplot)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Potential/figures/GL_potential_boxplot_",plot,".pdf"), width = 8, height = 6)
print(potential_load_boxplot)
dev.off()

potential_load_boxplot_wo_points <- ggplot(df_potential_load, aes(x = Group, y = Masked_GLm, fill = Group)) +
  #geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) + #points
  geom_boxplot(outlier.shape = NA, alpha = 0.6, color = "black", size = 0.6) + 
  scale_fill_manual(values = c("darkred", "darkgreen")) +
  scale_color_manual(values = c("darkred", "darkgreen")) + 
  labs(
    title = paste0("Boxplot Plot of Potential Genetic Load estimates from Provean ", plot),
    #x = "Group-age",
    y = "Potential Genetic Load (Provean)"
  ) +
  coord_cartesian(ylim = c(0.10, 0.22)) +  
  scale_y_continuous(breaks = seq(0.10, 0.22, by = 0.02)) +  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

plot(potential_load_boxplot_wo_points)

pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Potential/figures/GL_potential_boxplot_wopoints_",plot,".pdf"), width = 8, height = 6)
print(potential_load_boxplot_wo_points)
dev.off()
}

print(potential_load_difference_across_adults_juv_Lacanau)
print(potential_load_difference_across_adults_juv_Tocchi)
```

## Raw number of deleterious mutations in heterozygous

We computed the potential load without normalising by the number of neutral heterozygous sites because comparisons were performed among individuals within the same population, which are expected to share a common demographic history and similar levels of neutral genetic variation.

Final figure
```{r good colors figure}
list_plots <- c("Lacanau","Tocchi")

#Define color lists
color_list <- list(
  Lacanau = c(
    Adult = "#3A6FB0",
    Juvenile = "#A7C7E7"
  ),
  Tocchi = c(
    Adult = "#C49A00",
    Juvenile = "#FFD700"
  )
)
 
for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  df_potential_load <- get(paste0("Df_load_total_realised_",plot))
  
  t_test_result <- t.test(Heterozygous_deleterious_counts ~ Group, data = df_potential_load)

    assign(paste0("raw_potential_load_difference_across_adults_juv_",plot),t_test_result)
  
  #Select correct color palette
  col_pal <- color_list[[plot]]
  
  #Boxplot
  raw_potential_load_boxplot <- ggplot(df_potential_load, 
                                       aes(x = Group, 
                                           y = Heterozygous_deleterious_counts, 
                                           fill = Group)) +
    geom_jitter(aes(color = Group), width = 0.2, alpha = 0.6, size = 1.5, show.legend = FALSE) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = col_pal) +
    scale_color_manual(values = col_pal) +
    labs(
      title = paste0("Boxplot of Potential Raw Genetic Load (Provean) - ", plot),
      y = "Number of deleterious heterozygous mutations"
    ) +
    coord_cartesian(ylim = c(20, 140)) +  
  scale_y_continuous(breaks = seq(20, 140, by = 20)) +  
    theme_bw() +
theme(
  panel.grid.minor.y = element_blank(), 
  panel.grid.minor.x = element_blank(),   
  panel.grid.major.y = element_line(color="grey80"),  
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)
  
  plot(raw_potential_load_boxplot)

  #Save plot
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Potential/figures/GL_raw_potential_boxplot_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_potential_load_boxplot)
  dev.off()
  
  #Boxplot without the points
  raw_potential_load_boxplot_wo_points <- ggplot(df_potential_load, 
                                                 aes(x = Group, 
                                                     y = Heterozygous_deleterious_counts, 
                                                     fill = Group)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = col_pal) +
    labs(
      title = paste0("Boxplot of Potential Raw Genetic Load (Provean) - ", plot),
      y = "Number of deleterious heterozygous mutations"
    ) +
    coord_cartesian(ylim = c(20, 140)) +   
  scale_y_continuous(breaks = seq(20, 140, by = 20)) + 
    theme_bw() +
theme(
  panel.grid.minor.y = element_blank(),
  panel.grid.minor.x = element_blank(),  
  panel.grid.major.y = element_line(color="grey90"),  
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)

  plot(raw_potential_load_boxplot_wo_points)
  
  #Save plot without points
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Potential/figures/GL_raw_potential_boxplot_wopoints_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_potential_load_boxplot_wo_points)
  dev.off()
}

#Print t-test results
print(raw_potential_load_difference_across_adults_juv_Lacanau)
print(raw_potential_load_difference_across_adults_juv_Tocchi)
```

Both populations in the same figure
```{r plot full legend potential load}
#Merge the two datasets and create a new combined category
df_all <- rbind(
  transform(Df_load_total_realised_Lacanau, population = "Lacanau"),
  transform(Df_load_total_realised_Tocchi, population = "Tocchi")
)

#Create combined factor for x-axis: Population + Age
df_all$pop_age <- factor(
  paste(df_all$population, df_all$Group, sep = "_"),
  levels = c("Lacanau_Adult","Lacanau_Juvenile",
             "Tocchi_Adult","Tocchi_Juvenile")
)

#Colors:
custom_colors <- c(
  "Lacanau_Adult"     = "#3A6FB0",   
   "Lacanau_Juvenile" = "#A7C7E7",   
  "Tocchi_Adult"      = "#C49A00",    
    "Tocchi_Juvenile"  = "#FFD700"  
)

legend_colors <- c(
  "Lacanau" = "#3A6FB0",
  "Tocchi"  = "#E6C300"
)

raw_potential_load_boxplot_wo_points_combined <- ggplot(df_all, aes(x = pop_age, y = Heterozygous_deleterious_counts, fill = pop_age)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, color = "black", size = 0.6) +
    scale_fill_manual(values = custom_colors) +
    labs(
      title = "Boxplot of Potential Raw Genetic Load (Provean) - ",
      y = "Number of deleterious heterozygous mutations"
    ) +
    coord_cartesian(ylim = c(20, 140)) +
    scale_y_continuous(breaks = seq(20, 140, by = 20)) +
    theme_bw() +
      theme_bw() +
theme(
  panel.grid.minor.y = element_blank(),   
  panel.grid.minor.x = element_blank(),   
  panel.grid.major.y = element_line(color="grey90"),  
  plot.title = element_text(hjust = 0.5),
  legend.position = "none",
  axis.text.x = element_text(angle = 0, hjust = 0.5),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12)
)

  plot(raw_potential_load_boxplot_wo_points_combined)
  
  #Save plot without points
  pdf(paste0("Results/Chapter3/Pinaster/Genetic_load_new_estimates/Comparison_adults_juveniles/Potential/figures/GL_raw_potential_boxplot_wopoints_both_plots_",plot,".pdf"),
      width = 8, height = 6)
  print(raw_potential_load_boxplot_wo_points_combined)
  dev.off()
```

```{r significant differences potential load raw values}
#one way anova to see if groups are differents
anova_result_potential_raw <- aov(Heterozygous_deleterious_counts ~ pop_age, data = df_all)
summary(anova_result_potential_raw)

tukey_result <- TukeyHSD(anova_result_potential_raw)
tukey_result

anova_result_pop_potential_raw <- aov(Heterozygous_deleterious_counts ~ population, data = df_all)
summary(anova_result_pop_potential_raw)

#Compute median per group
df_all %>% group_by(population) %>% summarise(median = median(Heterozygous_deleterious_counts))

#Correlation between potential and realised load
cor(df_all$Homozygous_deleterious_counts,df_all$Heterozygous_deleterious_counts)

df_all_lac <- df_all %>% filter(population=="Lacanau")

cor(df_all_lac$Homozygous_deleterious_counts,df_all_lac$Heterozygous_deleterious_counts)

df_all_toc <- df_all %>% filter(population=="Tocchi")

cor(df_all_toc$Homozygous_deleterious_counts,df_all_toc$Heterozygous_deleterious_counts)
```