---
title: "Historical_climate_Ppinaster_study"
author: "Thomas Francisco"
date: "2025-07-18"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true 
    collapsed: True
    toc_depth: 3
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/tfrancisco/Documents/Thesis/")

rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
```

```{r library, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr) 
library(tidyverse)
library(ggplot2)
library(corrplot)  
library(dismo)
library(raster)
library(vegan)
library(plyr)
library(writexl)
library(rasterize)
library(sf)
library(FactoMineR)
library(ggrepel)
library(terra)
```

# Historical climates

This script investigates the historical climate that was present in Lacanau and Tocchi to better understand the history of the potential adaptive signals observed, by extracting and computing the climatic data across the retained climatic predictors.

We downloaded climatic data for the last interglacial (LIG~120,000- 140,000 years BP) and mid Holocene (~6,000 years BP) from Worldclim v1.4 database. The extracted climatic data contained mininum and maximum monthly temperatures as well as monthly precipitation. 
Dismo package was used to compute the bioclimatic variables. 
Summer heat moisture (SHM) index was computed as the ratio between mean warmest month temperature (MWMT) and 100* mean summer (May to September) precipitation (MSP). 

The codes with a # should be run, but they took a long time, so they were not rerun to produce the HTML file.  

```{r meta data}
meta_data_vcf=read.table("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Individual/Individual_code_wo2.txt",h=T)

meta_data_pop <- read.csv("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Population/Population_data_wo2.csv",h=T,sep=";",dec=",")

meta_data_pop_final <- meta_data_pop %>%
    filter(!(Population_ID %in% c("COM", "FUEbis")))

meta_data_pop_final <- meta_data_pop_final[,-c(1,2,4,10)]
meta_data_pop_final <- meta_data_pop_final[order(meta_data_pop_final$Population_ID), ]
meta_data_vcf_final <- meta_data_vcf[meta_data_vcf$Population_ID %in% meta_data_pop_final$Population_ID,] 

missing_in_df2 <- setdiff(meta_data_vcf$Population_ID, meta_data_pop_final$Population_ID)

save(meta_data_vcf_final, file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Individual/meta_data_vcf_final.Rdata")
save(meta_data_pop_final, file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Population/meta_data_pop_final.Rdata")
```

## LIG climate from Worldclim v1.4

We downloaded the data (BIOCLIM, PREC, TMAX and TMIN) from https://www.worldclim.org/data/v1.4/paleo1.4.html#google_vignette (Otto-Bliesner et al., 2008) at 30 arc-seconds resolution

### Bioclimatic variables 

We can load the bioclim variables which are in BIL format:
```{r load LIG data}
#Create an R object where all the climatic rasters are.
ras.bio_LIG <- stack(list.files("Data/Climatic/Ancient_climate/LIG/bio/retained/", pattern = ".bil$", full.names = T)) 

#Make sure it's in the right order
names(ras.bio_LIG) <- unlist(strsplit(unlist(lapply(strsplit(list.files("Data/Climatic/Ancient_climate/LIG/bio/retained/", pattern = ".bil$", full.names = T), split = "Data/Climatic/Ancient_climate/LIG/bio/retained/"), function(x) x[2])), split = ".bil"))
```

We need to compute the SHM variable obtained by MWMT/ (MSP*100). We need the monthly data for temperature and precipitation. We can load the monthly data. 
```{r load monthly data LIG}
#Precipitation (12 monthly files)
prec <- rast(list.files("Data/Climatic/Ancient_climate/LIG/lig_30s_prec/", pattern = "lig_30s_prec_.*\\.bil$", full.names = TRUE))

#Tmin / Tmax
tmin <- rast(list.files("Data/Climatic/Ancient_climate/LIG/lig_30s_tmin/", pattern = "lig_30s_tmin_.*\\.bil$", full.names = TRUE))

tmax <- rast(list.files("Data/Climatic/Ancient_climate/LIG/lig_30s_tmax/", pattern = "lig_30s_tmax_.*\\.bil$", full.names = TRUE))
```

We restricted the rasters to a reduce extend to speed up the analyses
```{r crop the raster and put in right format LIG}
#We can reduce the raster extend
extend_reduce <- ext(-10, 20, 30, 50)

#Crop
tmin_cropped <- crop(tmin, extend_reduce)

#Values were multiplied by 10 in Worldclim
tmin_cropped <- tmin_cropped / 10
tmin_cropped_order <- subset(tmin_cropped, c(1,5:12,2:4))

#Crop + values
tmax_cropped <- crop(tmax, extend_reduce)
tmax_cropped <- tmax_cropped /10
tmax_cropped_order <- subset(tmax_cropped, c(1,5:12,2:4))

#Prec
prec_cropped <- crop(prec, extend_reduce)
prec_cropped_order <- subset(prec_cropped, c(1,5:12,2:4))
```

We computed the bioclim variable to compare to the bioclim variables of Woldclim, to investigate if the Worldclim temperature variables are multiplied by 10.
```{r compute bio variables with biovar LIG}
#prec_cropped_format_biovar <- stack(prec_cropped_order)
#tmin_cropped_format_biovar <- stack(tmin_cropped_order)
#tmax_cropped_format_biovar <- stack(tmax_cropped_order)

#bioclim_LIG <- dismo::biovars(prec_cropped_format_biovar,tmin_cropped_format_biovar,tmax_cropped_format_biovar)

#bioclim_LIG_reduce <- bioclim_LIG[[1,3,4,12,15]]

#The script is highly inefficient. Instead, we can save the rasters and only extract the points for which we want to compute the bioclim variables in the RDA script

#save(prec_cropped_format_biovar,file="Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/prec_cropped_format_biovar.Rdata")
#writeRaster(prec_cropped_format_biovar, "Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/prec_cropped_format_biovar.tif", overwrite=TRUE)
#save(tmin_cropped_format_biovar,file="Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmin_cropped_format_biovar.Rdata")
#writeRaster(tmin_cropped_format_biovar, "Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmin_cropped_format_biovar.tif", overwrite=TRUE)
#save(tmax_cropped_format_biovar,file="Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmax_cropped_format_biovar.Rdata")
#writeRaster(tmax_cropped_format_biovar, "Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmax_cropped_format_biovar.tif", overwrite=TRUE)
```
This was the case. 

### SHM variable

Compute the MWMT and MSP: 
```{r compute MWMT and MSP LIG}
#Monthly mean temperature
tmean <- (tmin_cropped_order + tmax_cropped_order) / 2

#Mean Warmest Month Temperature
MWMT <- app(tmean, max)
MSP <- app(prec_cropped_order[[5:9]], mean)
```

Compute SHM
```{r compute SHM LIG}
SHM <- MWMT / (MSP / 1000)

coords<- data.frame(apply(meta_data_pop_final[,c(5:4)], 2, as.numeric)) #We need to have longitude then latitude
test <- data.frame(meta_data_pop_final$Population_ID,meta_data_pop_final$Country ,raster::extract(SHM, coords))

SHM_LIG <- raster(SHM)

#save(SHM_LIG,file="Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/SHM_LIG.Rdata")

writeRaster(SHM_LIG, "Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/SHM_LIG.tif", overwrite=TRUE)
```

Merge SHM with bioclim rasters:
```{r merge rasters LIG}
#Change the format of the raster
SHM_r <- raster(SHM)
#rename raster
names(SHM_r) <- "SHM"

#Reduce extend of Bio var
ras.bio_LIG_crop <- crop(ras.bio_LIG, SHM_r)

raster_bio_LIG_SHM_Worlclim <- stack(ras.bio_LIG_crop, SHM_r)

names(raster_bio_LIG_SHM_Worlclim) <- c("bio1", "bio12", "bio15","bio3","bio4","SHM")
#save(raster_bio_LIG_SHM_Worlclim,file="Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/raster_bio_LIG_SHM_Worlclim.Rdata")
writeRaster(raster_bio_LIG_SHM_Worlclim, "Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/raster_bio_LIG_SHM_Worlclim.tif", overwrite=TRUE)
```

## Mid Holocene 

First, we downloaded the CHELSA 21K climatic data and took the 10,000 BP information at a 30 arc-second resolution. However, this file only contains bioclim varibles, so we could not compute the SHM.  
Instead, we investigated the climate at the Mid-Holocene (~6,000 years BP) using the data available on WorldClim v1.4. We extracted the climatic data predicted by five different models. 

```{r load future climatic data}
#List of the model for the loop
list_models<- c("MIROC_ESM","IPSL_CM5A_LR","MPI_ESM_P","MRI_CGCM3","CNRM_CM5") 

for(x in 1:length(list_models)){
  
  model <- list_models[x]
#Create an R object where all the climatic rasters are
ras.bio_MH6000 <- stack(list.files(paste0("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"/bio_30s/retained/"), pattern = ".tif$", full.names = T)) 

#Order of the variables
names(ras.bio_MH6000) <- unlist(strsplit(unlist(lapply(strsplit(list.files(paste0("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"/bio_30s/retained/"), pattern = ".tif$", full.names = T), split = paste0("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"/bio_30s/retained/")), function(x) x[2])), split = ".tif"))

assign(paste0("ras_bio_MH6000_",model), ras.bio_MH6000)
}
```

Monthly climatic data for tmin, tmax and prec
```{r load monthly data Mid Holocene}
list_models<- c("MIROC_ESM","IPSL_CM5A_LR","MPI_ESM_P","MRI_CGCM3","CNRM_CM5") 

for(x in 1:length(list_models)){
  
  model <- list_models[x]
  
#Precipitation (12 monthly files)
prec <- rast(list.files(file.path("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"pr_30s/"), pattern = "\\.tif$", full.names = TRUE))

assign(paste0("prec_MH6000_",model), prec)

#Tmin / Tmax
tmin <- rast(list.files(file.path("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"tn_30s/"), pattern = "\\.tif$", full.names = TRUE))

assign(paste0("tmin_MH6000_",model), tmin)

tmax <- rast(list.files(file.path("Data/Climatic/Ancient_climate/Mid_Holo_6000/",model,"tx_30s/"), pattern = "\\.tif$", full.names = TRUE))

assign(paste0("tmax_MH6000_",model), tmax)
}
```

Computation of both the bioclimatic and SHM variables:
```{r crop the raster and put in right format Mid Holocene}
#for(x in 1:length(list_models)){
  
 # model <- list_models[x]

#We can reduce the raster extend to go faster
#extend_reduce <- ext(-10, 20, 30, 50)

#tmin <- get(paste0("tmin_MH6000_",model))
#Crop
#tmin_cropped <- crop(tmin, extend_reduce)

#Values were multiplied by 10 in Worldclim
#tmin_cropped <- tmin_cropped / 10
#tmin_cropped_order <- subset(tmin_cropped, c(1,5:12,2:4))

#Crop + values
#tmax <- get(paste0("tmax_MH6000_",model))

#tmax_cropped <- crop(tmax, extend_reduce)
#tmax_cropped <- tmax_cropped /10
#tmax_cropped_order <- subset(tmax_cropped, c(1,5:12,2:4))

#Prec
#prec <- get(paste0("prec_MH6000_",model))

#prec_cropped <- crop(prec, extend_reduce)
#prec_cropped_order <- subset(prec_cropped, c(1,5:12,2:4))

#Biovar format
#prec_cropped_format_biovar_Mid_Holo_6000 <- stack(prec_cropped_order)

#tmin_cropped_format_biovar_Mid_Holo_6000 <- stack(tmin_cropped_order)

#tmax_cropped_format_biovar_Mid_Holo_6000 <- stack(tmax_cropped_order)

#writeRaster(prec_cropped_format_biovar_Mid_Holo_6000, paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/prec/prec_cropped_format_biovar_Mid_Holo_6000_",model,".tif"), overwrite=TRUE)

#writeRaster(tmin_cropped_format_biovar_Mid_Holo_6000, paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/tmin/tmin_cropped_format_biovar_Mid_Holo_6000_",model,".tif"), overwrite=TRUE)

#writeRaster(tmax_cropped_format_biovar_Mid_Holo_6000, paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/tmax/tmax_cropped_format_biovar_Mid_Holo_6000_",model,".tif"), overwrite=TRUE)

#tmean <- (tmin_cropped_order + tmax_cropped_order) / 2

#Mean Warmest Month Temperature
#MWMT <- terra::app(
 # tmean,
  #fun = function(x) {
   # if (all(is.na(x))) {
    #  return(NA)
    #} else {
    #  return(max(x, na.rm = TRUE))
    #}
#  }
#)

#MSP <- app(prec_cropped_order[[5:9]], mean)

#SHM <- MWMT / (MSP / 1000)

#coords<- data.frame(apply(meta_data_pop_final[,c(5:4)], 2, as.numeric)) #We need to have longitude then latitude
#test <- data.frame(meta_data_pop_final$Population_ID,meta_data_pop_final$Country ,raster::extract(SHM, coords))

#SHM_MH_6000 <- raster(SHM)

#obj_name <- paste0("SHM_MH_6000_", model)

#assign(obj_name, SHM_MH_6000)

#save(list = obj_name,file = paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/SHM/",obj_name,".Rdata"))
    
#writeRaster(SHM_MH_6000, paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/SHM/SHM_MH_6000_",model,".tif"), overwrite=TRUE)

#Change format of the raster
#SHM_r <- raster(SHM)

#Rename raster
#names(SHM_r) <- "SHM"

#ras.bio_MH6000 <- get(paste0("ras_bio_MH6000_",model))

#Reduce extent of Bio var
#ras.bio_MH6000_crop <- crop(ras.bio_MH6000, SHM_r)

#raster_bio_MH6000_SHM_Worlclim <- stack(ras.bio_MH6000_crop, SHM_r)
#names(raster_bio_MH6000_SHM_Worlclim) <- c("bio1", "bio12", "bio15","bio3","bio4","SHM")

#writeRaster(raster_bio_MH6000_SHM_Worlclim, paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/SHM/raster_bio_MH6000_SHM_Worlclim_",model,".tif"), overwrite=TRUE)
#}
```