---
title: "RDA_model_within-population_and_evolutionary_responses_Ppinaster_study"
author: "Thomas Francisco"
date: "2025-07-22"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true 
    collapsed: True
    toc_depth: 3
---
```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/tfrancisco/Documents/Thesis/")

rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
```

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(vegan)
library(dplyr)
library(robust)
library(qvalue)
library(rgdal)
library(sf)
library(terra)
library(ggplot2)
library(radiant.data)
library(textshape)
library(rnaturalearth)
library(scales)
library(raster)
library(corrplot)
library(writexl)
library(ggspatial)
library(vcfR)
library(stringr)
library(dismo)
library(grid)
```

This script investigates the within-population variability in genomic composition at climate-associated loci, as well as potential evolutionary responses to climate change across cohorts. 

First, we built a GEA model using the genomic data across the 82 populations covering the species range and the climatic conditions for each of these populations for the retained climatic predictors during the 1901-1950 time period (past). This model was used to model the relationship between climate associated loci and the climatic predictors, as well as to predict the optimal genomic composition for different time periods across Lacanau and Tocchi populations:  
- Historical (Mid-Holocene, approximately 6,000 years ago)  
- Present (1991-2020)  
- Future (2041-2070)  

Finally, we transformed the genomic data at climate-associated loci of each individual in these two populations to RDA values and plotted then in this space along with optimal predicted values. 
Student t-test were performed to compare the mean genomic composition at climate associated loci across cohorts to detect potential evolutionary responses. 

# Data

```{r meta data}
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Population/meta_data_pop_final.Rdata")
#Alphabetic order
meta_data_pop_order <- meta_data_pop_final[order(meta_data_pop_final$Population_ID),]
identical(meta_data_pop_order$Population_ID,meta_data_pop_final$Population_ID)
```

## Past/present and future climatic data 

```{r load climatic data}
#climatic/structure/IBD data
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/GEA/variance_partitioning/Climatic_data_RDA_pRDA.Rdata") 
```

Here we will also load the raster data not scaled but with the mean and sd values of the past climatic data to scaled the data together following Capblancq & Forester (2021).
```{r load climatic data raster format}
list_models_clim<- c("GFDL_ESM4","IPSL_CM6A_LR","MPI_ESM1_2_HR","MRI_ESM2_0","UKESM1_0_LL") 

#Load past and future raster data
climatic_variables_to_keep <-c("bio1", "bio12", "bio15","bio3","bio4","SHM") #Subset of retained clim var, see script processing clim data
list_raster <- c("past_climatic_data_raster")#,"present_climatic_data_raster","future_climatic_data_raster"

for(i in 1:length(list_models_clim)){
  
  model <- list_models_clim[i]

for(x in 1:length(list_raster)){
  name <- c("past")#,"present","future"
  name_clim <- name[x]#name final raster
  var <- list_raster[x]#name of the raster
  
  if(var == "future_climatic_data_raster"){
    var= paste0("future_climatic_data_raster_",model)
    
    load(paste0("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_offset/RDA/",var,".Rdata")) #Load data
  raster_clim <- get(var)[[climatic_variables_to_keep]]#Keep only var in clim_variables to keep
  names(raster_clim)=c("Annual_Tc","Annual_P","P_Seasonality","Isothermality","Tc_Seasonality","Summer_aridity") #Change the name to match the names of the scaled and center values
  assign(paste0("raster_",name_clim,"_clim_",model),raster_clim) 
  } else {
    load(paste0("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_offset/RDA/",var,".Rdata")) #Load data
  raster_clim <- get(var)[[climatic_variables_to_keep]]#Keep only var in clim_variables to keep
  names(raster_clim)=c("Annual_Tc","Annual_P","P_Seasonality","Isothermality","Tc_Seasonality","Summer_aridity") #Change the name to match the names of the scaled and center values
  assign(paste0("raster_",name_clim,"_clim"),raster_clim) 
   }
  }
}
```

We also need the scale and center values for the reference period (past): 
```{r scale and center values}
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Climatic_data/scale_env_value_cli.Rdata")
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Climatic_data/center_env_value_cli.Rdata")
```

```{r extract and scale the climatic data}
list_name <- c("past")#,"present","future"
coords <- data.frame(apply(meta_data_pop_order[,c(5:4)], 2, as.numeric))#Longitude then latitude

for(i in 1:length(list_models_clim)){
  
  model <- list_models_clim[i]

for(x in 1: length(list_name)){
  
  name <- list_name[x]
  
  if(name == "future"){
    
  raster_data <- get(paste0("raster_",name,"_clim_",model))
  
  clim_data <- raster::extract(raster_data, coords)
    clim_data_order <- data.frame(clim_data[,c(1,4,5,2,3,6)])#Put it in the same order as center_env and scale object
  #Standardisation of the environmental variables
  clim_data_scale <- as.data.frame(scale(clim_data_order, center=center_env_value_cli, scale=scale_env_value_cli))
  assign(paste0("clim_df_",name,"_scale_",model),clim_data_scale)
  } else {
    
   raster_data <- get(paste0("raster_",name,"_clim"))
  
  clim_data <- raster::extract(raster_data, coords)
    clim_data_order <- data.frame(clim_data[,c(1,4,5,2,3,6)])#Same order as well
  #Standardisation of the environmental variables
  clim_data_scale <- as.data.frame(scale(clim_data_order, center=center_env_value_cli, scale=scale_env_value_cli))
  
  assign(paste0("clim_df_",name,"_scale"),clim_data_scale) 
    }
  }
}
```

## Genomic data
        
The genomic data were corrected using a more stringent MAC correction for the 82 populations.
There was an issue with the genotypes in the VCF file. Therefore, we need to recompute the allele frequencies for the 82 populations so that they match the genotypes used in the Lacanau and Tocchi datasets.

```{r genomic data}
#genomic data
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/data_allelic_frequencies_82pop_pinaster_9636_1518.Rdata")

colnames(data_allelic_frequencies_82pop_pinaster_9636_1518) <- sub("_.*", "", colnames(data_allelic_frequencies_82pop_pinaster_9636_1518))
genomic_matrix <- data_allelic_frequencies_82pop_pinaster_9636_1518
```

### Outliers sets
      
We computed the GEA models using four different set of SNPs to visualise the impact of the SNP sets on the GEA model. 

#### Less conservative thresholds  

```{r load outliers LC}
#set of less conservative thresholds
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/outliers_set_LC_common.Rdata")
```

#### More conservative thresholds

```{r load outliers MC}
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/outliers_set_MC_common.Rdata")
```

#### all outliers

```{r load outliers all outliers}
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/all_outliers_set_LC_common.Rdata")
all_outliers_set_LC_common_unique <- all_outliers_set_LC_common %>% unique()
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/all_outliers_set_MC_common.Rdata")
all_outliers_set_MC_common_unique <- all_outliers_set_MC_common %>% unique()
```

We also need to load 45 SNPs that seem not to match between the two datasets and that should therefore be removed just in case. 

```{r snp to remove}
SNPs_to_remove_45 <- read.table("Results/all_species/Genetic_load/Issue_polarisation/Pinaster_secondVCF/snp_non_matching_REF_ALT_to_remove.txt",h=T)

#Reformate
SNPs_to_remove_45$x <- sub("-", "", SNPs_to_remove_45$x)
#Apply it to all outlier LC
all_outliers_set_LC_common_unique_final <- all_outliers_set_LC_common_unique[ ! all_outliers_set_LC_common_unique %in% SNPs_to_remove_45$x ]

#All MC
all_outliers_set_MC_common_unique_final <- all_outliers_set_MC_common_unique[ ! all_outliers_set_MC_common_unique %in% SNPs_to_remove_45$x ]

#LC
outliers_set_LC_common_final <- outliers_set_LC_common[ ! outliers_set_LC_common %in% SNPs_to_remove_45$x ]

#MC
outliers_set_MC_common_final <- outliers_set_MC_common[ ! outliers_set_MC_common %in% SNPs_to_remove_45$x ]
```

# Calculation of the GEA relationship

## Models

### Less conservative outlier set (LC)

RDA model built with all the climate-associated SNPs identified by at least two GEA methods and using the less conservative thresholds for identification: FDR 10%, top 5% and Bayes factor < 8; see explanations on the README file about the climate-associated SNPs identification.  
```{r RDA on LC outliers set}
RDA_outliers_LC <- rda(formula = genomic_matrix[outliers_set_LC_common_final] ~  Annual_Tc+Isothermality+Tc_Seasonality+Annual_P+P_Seasonality+Summer_aridity, data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(RDA_outliers_LC)
```

### More conservative outlier set (MC)

RDA model built with climate-associated SNPs identified by at least two GEA methods and using the more conservative thresholds for identification: FDR 5%, top 1% and Bayes factor < 10; see also explanations on the README file about the climate-associated SNPs identification. 
```{r RDA on MC outliers set}
RDA_outliers_MC <- rda(formula = genomic_matrix[outliers_set_MC_common_final] ~  Annual_Tc+Isothermality+Tc_Seasonality+Annual_P+P_Seasonality+Summer_aridity, data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(RDA_outliers_MC)
```

### All outliers set LC

RDA model built with all the climate-associated SNPs identified using the less conservative thresholds. 
```{r RDA on all_outliers set LC}
RDA_all_outliers_LC <- rda(formula = genomic_matrix[all_outliers_set_LC_common_unique_final] ~  Annual_Tc+Isothermality+Tc_Seasonality+Annual_P+P_Seasonality+Summer_aridity, data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(RDA_all_outliers_LC)
```

### All outliers set MC

RDA model built with all the climate-associated SNPs identified using the more conservative thresholds. 
```{r RDA on all_outliers set MC}
RDA_all_outliers_MC <- rda(formula = genomic_matrix[all_outliers_set_MC_common_unique_final] ~  Annual_Tc+Isothermality+Tc_Seasonality+Annual_P+P_Seasonality+Summer_aridity, data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(RDA_all_outliers_MC)
```

We can save the models to used them later for comparison:
```{r save rda models}
list_models <- c("RDA_outliers_LC","RDA_outliers_MC","RDA_all_outliers_LC","RDA_all_outliers_MC")

for(x in 1:length(list_models)){
  
  name_model<- list_models[x]
  file_save<- get(name_model)
  name <- paste0("model_GEA_",name_model)
  assign(name,file_save)
  save(list = paste0("model_GEA_",name_model),file=paste0("Results/Chapter3/Pinaster/GEA_model_pinaster/Genomic_offset/RDA/data/models/",name,".Rdata"))
}
```

All further analysis will be performed on the different climate-associated sets to investigate whether we observe the same patterns regardless of the SNP sets: 
```{r list set}
list_outliers_set <- c("outliers_set_LC_common_final","outliers_set_MC_common_final","all_outliers_set_LC_common_unique_final","all_outliers_set_MC_common_unique_final")
list_RDA <- c("RDA_outliers_LC","RDA_outliers_MC","RDA_all_outliers_LC","RDA_all_outliers_MC")
```

## Graphical visualisation

We can plot the explained variance along the RDA axes and the climatic variables: 
```{r screeplot outliers LC, eval=F, echo=T}
for(x in 1: length(list_RDA)){
  RDA <- get(list_RDA[x])
  name <- c("outliers_LC","Outliers_MC","all_outliers_LC","all_outliers_MC")
  #Screeplot
plot(RDA$CCA$eig, option="screeplot")

#In histogram
screeplot_RDA <- screeplot(RDA,main = paste0("Hist screeplot ",name[x]))

#Explained variance along each RDA axis
explain_variance <-RDA$CCA$eig*100/sum(RDA$CCA$eig)

assign(paste0("explain_variance",name[x]),explain_variance)

  print(get(paste0("explain_variance",name[x])))
#Save
pdf(paste0("Results/Chapter3/Pinaster/GEA_model_pinaster/Genomic_offset/RDA/figures/screeplot/Screeplot_RDA_",name[x],".pdf"));screeplot(RDA,main = paste0("Hist screeplot ",name[x]));dev.off()
}
```

```{r graphic rmarkdown, echo=FALSE}
list_RDA_rmarkdown <- c("RDA_all_outliers_MC")
for(x in 1: length(list_RDA_rmarkdown)){
  RDA <- get(list_RDA_rmarkdown[x])
  name <- c("outliers_Lc")
  #Screeplot
plot(RDA$CCA$eig, option="screeplot")

#In histogram
screeplot_RDA <- screeplot(RDA,main = paste0("Hist screeplot ",name[x]))

#Explained variance along each RDA axis
explain_variance <-RDA$CCA$eig*100/sum(RDA$CCA$eig)

assign(paste0("explain_variance",name[x]),explain_variance)

  print(get(paste0("explain_variance",name[x])))
}
```

We can see that the first two RDA axes appear to explain most of the genetic variation across populations. We will use these two axes for further analysis of each SNP set. 

We can also plot the loci and climatic variables in the RDA space to visualise their relationships:
```{r plot biplot loci,eval=F, echo=T}
for(x in 1:length(list_RDA)){
  
  RDA <- get(list_RDA[x])
  name <- c("outliers_LC","Outliers_MC","all_outliers_LC","all_outliers_MC")
  
  score_loci <- as.data.frame(scores(RDA, choices=c(1:2), display="species", scaling="none"))
score_climatic_var <- as.data.frame(scores(RDA, choices=c(1:2), display="bp"))
explained_variance_round <- round(RDA$CCA$eig*100/sum(RDA$CCA$eig),digits=1)

#Biplot with SNPs and climatic variables along the two first RDA axis. 
 biplot_outliers_RDA<- ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = score_loci, aes(x=RDA1*15, y=RDA2*15,color="locus"), size = 1.4) +
  geom_segment(data = score_climatic_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 2.5) +
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
   ggtitle(paste0("Biplot RDA ",name[x])) +
  guides(color=guide_legend(title="Locus type")) +
  scale_color_manual(values ="#F9A242FF") +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
 
 print(biplot_outliers_RDA)
 
#Save plots
pdf(paste0("Results/Chapter3/Pinaster/GEA_model_pinaster/Genomic_offset/RDA/figures/biplot/biplot_outliers_RDA_",name[x],".pdf"));print(biplot_outliers_RDA);dev.off()
}
```

```{r graph rmarkdown_2, echo=FALSE}
list_RDA_rmarkdown<- c("RDA_all_outliers_MC")

for(x in 1:length(list_RDA_rmarkdown)){
  
  RDA <- get(list_RDA_rmarkdown[x])
  name <- c("all","outliers_Lc")
  
  score_loci <- as.data.frame(scores(RDA, choices=c(1:2), display="species", scaling="none"))
score_climatic_var <- as.data.frame(scores(RDA, choices=c(1:2), display="bp"))
explained_variance_round <- round(RDA$CCA$eig*100/sum(RDA$CCA$eig),digits=1)

#Biplot with SNPs and climatic variables along the two first RDA axis. 
 biplot_outliers_RDA<- ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = score_loci, aes(x=RDA1*15, y=RDA2*15,color="locus"), size = 1.4) +
  geom_segment(data = score_climatic_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 2.5) +
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
   ggtitle(paste0("Biplot RDA ",name[x])) +
  guides(color=guide_legend(title="Locus type")) +
  scale_color_manual(values ="#F9A242FF") +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
 
 print(biplot_outliers_RDA)
}
```

We can plot the populations in this space to see if their distribution matches the expected climatic pattern. Here, the population positions represent observed values rather than predicted ones.
```{r biplot pop on the outliers RDA space,eval=F, echo=T}
for(x in 1:length(list_RDA)){
  RDA <- get(list_RDA[x])
  name <- c("outliers_LC","Outliers_MC","all_outliers_LC","all_outliers_MC")
  
#Score along the 2 first RDA axis
score_climatic_var <- as.data.frame(scores(RDA, choices=c(1:2), display="bp"))
#Score_population <- data.frame(RDA$CCA$u[,c(1,2)]) #predicted population scores
observed_scores_populations <- as.data.frame(scores(RDA,choices=c(1,2), display = "sites",scaling="none"))

#Explained variance
explained_variance_round <- round(RDA$CCA$eig*100/sum(RDA$CCA$eig),digits=1)

#Merge for country info
Score_population_bis <- rownames_to_column(observed_scores_populations,"Population_ID")
score_with_country_info <- merge(Score_population_bis,meta_data_pop_order[,c(1,3)],"Population_ID")
score_with_country_info$Country <- as.factor(score_with_country_info$Country)

group_palette <- c("ALG"="orangered3","ESP"= "darkorchid3", "FRA"="navyblue", "MAR"="turquoise2", "PRT"="orange", "TUN"="darkgreen")

#Biplot with populations and climatic variables along the 2 first RDA axis
biplot_populations <- ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_point(data = score_with_country_info, aes(x = RDA1 * 3, y = RDA2 * 3, colour = Country), size = 2, alpha = 0.8) +
  geom_segment(data = score_climatic_var, aes(xend = RDA1, yend = RDA2, x = 0, y = 0), colour = "black", size = 0.15, linetype = 1, arrow = arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 2.5)+
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
  ggtitle(paste0("Biplot RDA Populations ",name[x])) +
  scale_color_manual(name = "Countries", values = group_palette, labels = levels(score_with_country_info$Country)) +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))+
  labs(color = "Country")

 print(biplot_populations)
 
#Save plots
pdf(paste0("Results/Chapter3/Pinaster/GEA_model_pinaster/Genomic_offset/RDA/figures/biplot/biplot_population_RDA_",name[x],".pdf"));print(biplot_populations);dev.off()
}
```

```{r graph rmarkdown 3, echo=FALSE}
for(x in 1:length(list_RDA_rmarkdown)){
  RDA <- get(list_RDA_rmarkdown[x])
  name <- c("all_outliers_MC")
  
#score along the 2 first RDA axis
score_climatic_var <- as.data.frame(scores(RDA, choices=c(1:2), display="bp"))
#Score_population <- data.frame(RDA$CCA$u[,c(1,2)]) #predicted population scores
observed_scores_populations <- as.data.frame(scores(RDA,choices=c(1,2), display = "sites",scaling="none"))

#explained variance
explained_variance_round <- round(RDA$CCA$eig*100/sum(RDA$CCA$eig),digits=1)

#merge for country info
Score_population_bis <- rownames_to_column(observed_scores_populations,"Population_ID")
score_with_country_info <- merge(Score_population_bis,meta_data_pop_order[,c(1,3)],"Population_ID")
score_with_country_info$Country <- as.factor(score_with_country_info$Country)


   group_palette <- c("ALG"="orangered3","ESP"= "darkorchid3", "FRA"="navyblue", "ITA"="red", "MAR"="turquoise2", "PRT"="orange", "TUN"="darkgreen")


#Biplot with populations and climatic variables along the 2 first RDA axis
biplot_populations <- ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_point(data = score_with_country_info, aes(x = RDA1 * 3, y = RDA2 * 3, colour = Country), size = 2, alpha = 0.8) +
  geom_segment(data = score_climatic_var, aes(xend = RDA1, yend = RDA2, x = 0, y = 0), colour = "black", size = 0.15, linetype = 1, arrow = arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 2.5)+
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
  ggtitle(paste0("Biplot RDA Populations ",name[x])) +
  scale_color_manual(name = "Countries", values = group_palette, labels = levels(score_with_country_info$Country)) +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))+
  labs(color = "Country")

 print(biplot_populations)
}
```

# Historical climatic data 

This part was added after the results were obtained, enabling the historical climate (LIG and Mid-Holocene) to be extracted for the Lacanau and Tocchi populations. This makes it possible to obtain the optimal genomic composition predicted by the model for these historical periods.
```{r coordinates LAC and TOC}
meta_data_plots <- read.csv("Data/Chapter3/Pinus_pinaster/CGs/Metadata_Lac-Toc.csv",h=T,sep=";",dec=",")
```

First, we need to load the climatic variables and compute the BIO variables, and then add the SHM:
```{r load rasters LIG}
#LIG
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/prec_cropped_format_biovar.Rdata")
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmin_cropped_format_biovar.Rdata")
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmax_cropped_format_biovar.Rdata")
#prec_cropped_format_biovar_LIG <- prec_cropped_format_biovar
#tmin_cropped_format_biovar_LIG <- tmin_cropped_format_biovar
#tmax_cropped_format_biovar_LIG <- tmax_cropped_format_biovar

#prec_cropped_format_biovar_LIG <- stack("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/prec_cropped_format_biovar.tif")
#tmin_cropped_format_biovar_LIG <- stack("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmin_cropped_format_biovar.tif")
#tmax_cropped_format_biovar_LIG <- stack("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/tmax_cropped_format_biovar.tif")

#load("Data/Chapter3/Pinus_pinaster/Historical_climate/LIG/SHM_LIG.Rdata")

#MH6000
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000/prec_cropped_format_biovar_Mid_Holo_6000.Rdata")
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000/tmin_cropped_format_biovar_Mid_Holo_6000.Rdata")
#load("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000/tmax_cropped_format_biovar_Mid_Holo_6000.Rdata")

list_models<- c("MIROC_ESM","IPSL_CM5A_LR","MPI_ESM_P","MRI_CGCM3","CNRM_CM5") 

for(x in 1:length(list_models)){
  
  model <- list_models[x]

prec_cropped_format_biovar_Mid_Holo_6000 <- stack(paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/prec/prec_cropped_format_biovar_Mid_Holo_6000_",model,".tif"))
assign(paste0("prec_cropped_format_biovar_Mid_Holo_6000_",model),prec_cropped_format_biovar_Mid_Holo_6000)

tmin_cropped_format_biovar_Mid_Holo_6000 <- stack(paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/tmin/tmin_cropped_format_biovar_Mid_Holo_6000_",model,".tif"))
assign(paste0("tmin_cropped_format_biovar_Mid_Holo_6000_",model),tmin_cropped_format_biovar_Mid_Holo_6000)

tmax_cropped_format_biovar_Mid_Holo_6000 <- stack(paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/tmax/tmax_cropped_format_biovar_Mid_Holo_6000_",model,".tif"))
assign(paste0("tmax_cropped_format_biovar_Mid_Holo_6000_",model),tmax_cropped_format_biovar_Mid_Holo_6000)

load(paste0("Data/Chapter3/Pinus_pinaster/Historical_climate/Mid_Holo_6000_models/SHM/SHM_MH_6000_",model,".Rdata"))
}
```

Extract the climate for the two populations:
```{r point extraction coordonnees}
list_name <- c("Mid_Holo_6000_MIROC_ESM","Mid_Holo_6000_IPSL_CM5A_LR","Mid_Holo_6000_MPI_ESM_P","Mid_Holo_6000_MRI_CGCM3","Mid_Holo_6000_CNRM_CM5")#"LIG",
list_name_bis <- c("MH_6000_MIROC_ESM","MH_6000_IPSL_CM5A_LR","MH_6000_MPI_ESM_P","MH_6000_MRI_CGCM3","MH_6000_CNRM_CM5")#"LIG",
coords <- data.frame(apply(meta_data_plots[,c(2:3)], 2, as.numeric))#Longitude then latitude

for(x in 1: length(list_name)){
  
  name <- list_name[x]
  name_bis <- list_name_bis[x]
  
  prec_raster <- get(paste0("prec_cropped_format_biovar_",name))
  tmin_raster <- get(paste0("tmin_cropped_format_biovar_",name))
  tmax_raster <- get(paste0("tmax_cropped_format_biovar_",name))
  
  prec_data <- raster::extract(prec_raster, coords)
  tmin_data <- raster::extract(tmin_raster, coords)
  tmax_data <- raster::extract(tmax_raster, coords)

  bioclim_var <- biovars(prec_data,tmin_data,tmax_data)
  
  #SHM
  SHM_data <- get(paste0("SHM_",name_bis))
  
  SHM_extraction <- raster::extract(SHM_data, coords)
  
  df_clim_data <- data.frame(bioclim_var[,c(1,3,4,12,15)],SHM=SHM_extraction)
  colnames(df_clim_data) <- c("Annual_Tc","Isothermality","Tc_Seasonality","Annual_P","P_Seasonality","Summer_aridity")
  
  assign(paste0("clim_df_",name,"_nonscaled_plots"),df_clim_data)
  
  #Scaled
  clim_data_scale <- as.data.frame(scale(df_clim_data, center=center_env_value_cli, scale=scale_env_value_cli))
  assign(paste0("clim_df_",name,"_scale_plots"),clim_data_scale)
}
```

# Climatic data Lacanau and Tocchi for 1901-1950, 1991-2020 and 2041-2070 time periods

```{r load climatic data raster format LAC TOC}
list_models_clim<- c("GFDL_ESM4","IPSL_CM6A_LR","MPI_ESM1_2_HR","MRI_ESM2_0","UKESM1_0_LL") 

#Load past and future raster data
climatic_variables_to_keep <-c("bio1", "bio12", "bio15","bio3","bio4","SHM") #Subset of retained climatic variables, see script processing clim data
list_raster <- c("past_climatic_data_raster","present_climatic_data_raster","future_climatic_data_raster")

for(i in 1:length(list_models_clim)){
  
  model <- list_models_clim[i]

for(x in 1:length(list_raster)){
  name <- c("past","present","future")
  name_clim <- name[x]
  var <- list_raster[x]
  
  if(var == "future_climatic_data_raster"){
    var= paste0("future_climatic_data_raster_",model)
    
    load(paste0("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_offset/RDA/",var,".Rdata")) #Load data
  raster_clim <- get(var)[[climatic_variables_to_keep]]#Keep only the retained variables
  names(raster_clim)=c("Annual_Tc","Annual_P","P_Seasonality","Isothermality","Tc_Seasonality","Summer_aridity") #Change the name to match the names of the scaled and center values
  
  assign(paste0("raster_",name_clim,"_clim_",model),raster_clim) 
  
  } else {
    
    load(paste0("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_offset/RDA/",var,".Rdata")) #load data
  raster_clim <- get(var)[[climatic_variables_to_keep]]
  names(raster_clim)=c("Annual_Tc","Annual_P","P_Seasonality","Isothermality","Tc_Seasonality","Summer_aridity")
  
  assign(paste0("raster_",name_clim,"_clim"),raster_clim) 
    }
  }
}
```

```{r extract and scale the climatic data LAC and TOC}
list_name <- c("past","present","future")
coords <- data.frame(apply(meta_data_plots[,c(2:3)], 2, as.numeric))#Longitude then latitude

for(i in 1:length(list_models_clim)){
  
  model <- list_models_clim[i]

for(x in 1: length(list_name)){
  
  name <- list_name[x]
  
  if(name == "future"){
    
  raster_data <- get(paste0("raster_",name,"_clim_",model))
  
  clim_data <- raster::extract(raster_data, coords)
    clim_data_order <- data.frame(clim_data[,c(1,4,5,2,3,6)])#Same order as center_env and scale object
  # Standardisation of the environmental variables
  clim_data_scale <- as.data.frame(scale(clim_data_order, center=center_env_value_cli, scale=scale_env_value_cli))
  
  assign(paste0("clim_df_",name,"_nonscaled_plots_",model),clim_data_order)
  assign(paste0("clim_df_",name,"_scale_plots_",model),clim_data_scale)
  
  } else {
    
   raster_data <- get(paste0("raster_",name,"_clim"))

  clim_data <- raster::extract(raster_data, coords)
    clim_data_order <- data.frame(clim_data[,c(1,4,5,2,3,6)])
  clim_data_scale <- as.data.frame(scale(clim_data_order, center=center_env_value_cli, scale=scale_env_value_cli))
  
  assign(paste0("clim_df_",name,"_nonscaled_plots"),clim_data_order)
  assign(paste0("clim_df_",name,"_scale_plots"),clim_data_scale) 
    }
  }
}

future_nonscaled_list  <- lapply(list_models_clim, function(model) {
  get(paste0("clim_df_future_nonscaled_plots_", model))
})

future_scaled_list  <- lapply(list_models_clim, function(model) {
  get(paste0("clim_df_future_scale_plots_", model))
})

clim_df_future_mean_nonscaled_plots <- Reduce("+", future_nonscaled_list) / length(future_nonscaled_list)
clim_df_future_mean_scale_plots <- Reduce("+", future_scaled_list) / length(future_scaled_list)
```

# Compute the delta of climate between past and present for each climatic variable

This section computes the difference between the historical (mid-Holocene), present and future climates in order to gain insight into climatic variability over time and between populations. If one population has greater climatic differences than the other, this could suggest higher selective pressure, which may be translated in greater differences between adults and juveniles
```{r delta clim var historical}
list_df <- c("nonscaled","scale")
#list_plots <- c("Lacanau","Tocchi")

#for(x in 1:length(list_df)){
  
 # scaled_or_not <- list_df[x]
  
  #clim_df_Mid_Holo_6000 <- get(paste0("clim_df_Mid_Holo_6000_",scaled_or_not,"_plots"))
  #clim_df_Mid_Holo_6000$Population <- c("Lacanau","Tocchi")
  
  #clim_df_present <- get(paste0("clim_df_present_",scaled_or_not,"_plots"))
  #clim_df_present$Population <- c("Lacanau","Tocchi")

  #df_Lacanau_merged <- rbind(clim_df_Mid_Holo_6000[1,],clim_df_present[1,])
  #df_Tocchi_merged <- rbind(clim_df_Mid_Holo_6000[2,],clim_df_present[2,])

  #Delta_Lacanau <- df_Lacanau_merged[2,c(1:6) ] - df_Lacanau_merged[1,c(1:6) ]
#Delta_Tocchi  <- df_Tocchi_merged[2,c(1:6) ]  - df_Tocchi_merged[1,c(1:6) ]
  
#Delta_tot_merge <- data.frame(Population=c("Lacanau","Tocchi"),rbind(Delta_Lacanau,Delta_Tocchi))

#assign(paste0("Delta_tot_merge_historical_present_",scaled_or_not),Delta_tot_merge)
#}
```

```{r delta clim var present}
list_df <- c("nonscaled","scale")
#list_plots <- c("Lacanau","Tocchi")

for(x in 1:length(list_df)){
  
  scaled_or_not <- list_df[x]
  
  clim_df_present <- get(paste0("clim_df_present_",scaled_or_not,"_plots"))
  clim_df_present$Population <- c("Lacanau","Tocchi")
  
  clim_df_future_mean <- get(paste0("clim_df_future_mean_",scaled_or_not,"_plots"))
  clim_df_future_mean$Population <- c("Lacanau","Tocchi")

  df_Lacanau_merged <- rbind(clim_df_present[1,],clim_df_future_mean[1,])
  df_Tocchi_merged <- rbind(clim_df_present[2,],clim_df_future_mean[2,])

  Delta_Lacanau <- df_Lacanau_merged[2,c(1:6) ] - df_Lacanau_merged[1,c(1:6) ]
Delta_Tocchi  <- df_Tocchi_merged[2,c(1:6) ]  - df_Tocchi_merged[1,c(1:6) ]
  
Delta_tot_merge <- data.frame(Population=c("Lacanau","Tocchi"),rbind(Delta_Lacanau,Delta_Tocchi))

assign(paste0("Delta_tot_merge_present_future_",scaled_or_not),Delta_tot_merge)
}
```

# Compute the optimal genomic composition

The optimal genomic compositions were predicted for all the mentioned periods (LIG, mid-Holo, past, present and future) using the RDA model, which was built using data from 82 populations across the range. 
```{r predictions_adaptive_values}
list_models<- c("MIROC_ESM","IPSL_CM5A_LR","MPI_ESM_P","MRI_CGCM3","CNRM_CM5") 

list_rda <- c("RDA_outliers_LC","RDA_outliers_MC","RDA_all_outliers_LC",'RDA_all_outliers_MC')
list_period <- c("Mid_Holo_6000_MIROC_ESM","Mid_Holo_6000_IPSL_CM5A_LR","Mid_Holo_6000_MPI_ESM_P","Mid_Holo_6000_MRI_CGCM3","Mid_Holo_6000_CNRM_CM5","past","present","future")#"LIG",
list_set <- c("LC","MC","all_outliers_LC","all_outliers_MC")

K=2 #Number of rda axes retained

for(i in 1: length(list_rda)){
  
  rda_model <- get(list_rda[i])
  set <- list_set[i]
  
  for(x in 1: length(list_period)){
    
  name <- list_period[x]
  
  if(name == "future"){
  
  for(a in 1:length(list_models_clim)){
  
  models_clim <- list_models_clim[a]
  clim <- get(paste0("clim_df_",name,"_scale_plots_",models_clim))
  predictions <- data.frame(Population=meta_data_plots$ID,predict(rda_model,clim, type = "lc", rank = K, scaling = "none"))
  assign(paste0("adaptive_values_",set,"_",name,"_",models_clim),predictions)
  }
  
  }else{
    
    clim <- get(paste0("clim_df_",name,"_scale_plots"))
  #Model
  predictions <- data.frame(Population=meta_data_plots$ID,predict(rda_model,clim, type = "lc", rank = K, scaling = "none"))
  
  assign(paste0("adaptive_values_",set,"_",name),predictions)
  
    }
  }
}
#Mean value for future across climatic models

K <- 2#Number of RDA axes

#Loop over SNP sets
for (set in list_set) {
  
  model_names <- list_models_clim
  
  #Initialise list to hold renamed data
  future_preds_list <- list()
  
  for (model in model_names) {
    
    #Get the object
    df <- get(paste0("adaptive_values_", set, "_future_", model))
    
    #Rename RDA columns to avoid name conflict
    new_colnames <- c("Population", paste0("RDA", 1:K, "_", model))
    colnames(df)[1:(K+1)] <- new_colnames  #Assumes population is first
    
    future_preds_list[[model]] <- df
  }
  
  #Merge all by population
  future_preds_merged <- Reduce(function(x, y) merge(x, y, by = "Population"), future_preds_list)
  
  #Compute mean RDA scores across models
  future_mean_df <- data.frame(
    Population = future_preds_merged$Population,
    sapply(1:K, function(k) {
      #Get all RDA_k_* columns
      cols_k <- grep(paste0("RDA", k, "_"), colnames(future_preds_merged), value = TRUE)
      rowMeans(future_preds_merged[, cols_k], na.rm = TRUE)
    })
  )
  
  colnames(future_mean_df)[-1] <- paste0("RDA", 1:K)
  
  #Save final result
  obj_name <- paste0("adaptive_values_", set, "_future_mean")
  assign(obj_name, future_mean_df)
  
  #Store the object name for later saving
  save(list = obj_name, file = paste0("Results/Chapter3/Pinaster/RDA_values/adaptive_values/adaptive_values_", set, "_future_mean.Rdata"))
}
```

Optimal values including mid-Holocene:
```{r optimal values with mid holocene}
list_models <- c("MIROC_ESM","IPSL_CM5A_LR","MPI_ESM_P","MRI_CGCM3","CNRM_CM5") 

list_rda <- c("RDA_outliers_LC","RDA_outliers_MC","RDA_all_outliers_LC","RDA_all_outliers_MC")
list_period <- c("Mid_Holo_6000","past","present","future")#"LIG",
list_set <- c("LC","MC","all_outliers_LC","all_outliers_MC")

K <- 2

for (i in 1:length(list_rda)) {
  
  rda_model <- get(list_rda[i])
  set <- list_set[i]
  
  for (period in list_period) {
    
    if (period == "future") {
      
      for (model in list_models_clim) {
        
        clim <- get(paste0("clim_df_", period, "_scale_plots_", model))
        
        predictions <- data.frame(
          Population = meta_data_plots$ID,
          predict(rda_model, clim, type = "lc", rank = K, scaling = "none")
        )
        
        assign(
          paste0("adaptive_values_", set, "_", period, "_", model),
          predictions
        )
      }
      
    } else if (period == "Mid_Holo_6000") {
      
      for (model in list_models) {
        
        clim <- get(paste0("clim_df_", period, "_", model, "_scale_plots"))
        
        predictions <- data.frame(
          Population = meta_data_plots$ID,
          predict(rda_model, clim, type = "lc", rank = K, scaling = "none")
        )
        
        assign(
          paste0("adaptive_values_", set, "_", period, "_", model),
          predictions
        )
      }
      
    } else {
      
      clim <- get(paste0("clim_df_", period, "_scale_plots"))
      
      predictions <- data.frame(
        Population = meta_data_plots$ID,
        predict(rda_model, clim, type = "lc", rank = K, scaling = "none")
      )
      
      assign(
        paste0("adaptive_values_", set, "_", period),
        predictions
      )
    }
  }
}


for (set in list_set) {
  
  #FUTURE mean
  {
    preds_list <- list()
    
    for (model in list_models_clim) {
      
      df <- get(paste0("adaptive_values_", set, "_future_", model))
      
      colnames(df)[1:(K+1)] <- c("Population", paste0("RDA", 1:K, "_", model))
      
      preds_list[[model]] <- df
    }
    
    merged <- Reduce(function(x,y) merge(x,y,by="Population"), preds_list)
    
    future_mean_df <- data.frame(
      Population = merged$Population,
      sapply(1:K, function(k) {
        cols <- grep(paste0("RDA",k,"_"), colnames(merged), value=TRUE)
        rowMeans(merged[,cols], na.rm=TRUE)
      })
    )
    
    colnames(future_mean_df)[-1] <- paste0("RDA",1:K)
    
    obj_name <- paste0("adaptive_values_", set, "_future_mean")
    assign(obj_name, future_mean_df)
    
    save(list=obj_name,
         file=paste0("Results/Chapter3/Pinaster/RDA_values/adaptive_values/",obj_name,".Rdata"))
  }
  
  #MID HOLOCENE MEAN
  {
    preds_list <- list()
    
    for (model in list_models) {
      
      df <- get(paste0("adaptive_values_", set, "_Mid_Holo_6000_", model))
      
      colnames(df)[1:(K+1)] <- c("Population", paste0("RDA", 1:K, "_", model))
      
      preds_list[[model]] <- df
    }
    
    merged <- Reduce(function(x,y) merge(x,y,by="Population"), preds_list)
    
    mh_mean_df <- data.frame(
      Population = merged$Population,
      sapply(1:K, function(k) {
        cols <- grep(paste0("RDA",k,"_"), colnames(merged), value=TRUE)
        rowMeans(merged[,cols], na.rm=TRUE)
      })
    )
    
    colnames(mh_mean_df)[-1] <- paste0("RDA",1:K)
    
    obj_name <- paste0("adaptive_values_", set, "_Mid_Holo_6000_mean")
    assign(obj_name, mh_mean_df)
    
    save(list=obj_name,
         file=paste0("Results/Chapter3/Pinaster/RDA_values/adaptive_values/",obj_name,".Rdata"))
  }
}
```

## Genomic data Lacanau and Tocchi individuals

We need to calculate RDA scores for each individual in the Lacanau and Tocchi populations, based on their genomic composition at loci associated with climate. First, we need to format their genomic data.

Load genomic data:
```{r load genomic data plots}
load("Data/Chapter3/Pinus_pinaster/genomic_data_newpop/Adults_offsprings_merge_df/data_Lacanau_tot.Rdata")
colnames(data_Lacanau_tot) <- gsub("\\.", "", colnames(
  data_Lacanau_tot))

load("Data/Chapter3/Pinus_pinaster/genomic_data_newpop/Adults_offsprings_merge_df/data_Tocchi_tot.Rdata")
colnames(data_Tocchi_tot) <- gsub("\\.", "", colnames(data_Tocchi_tot))
```

We need to encode the SNPs in allelic frequency. To do this, we need to retrieve from the VCF which allele is the REF and which is the ALT, in order to match the REF and ALT used to build the GEA model. We should indeed use the same REF and ALT, otherwise, we will compare SNPs between the 82 range-wide populations from the model and the Lacanau and Tocchi individuals, which should not be compared as the information is displayed differently. 
We load the VCF with good genotypes (see the 'Correcting_genotypes_pinaster_VCF' script).
```{r load vcf corrected REF and ALT}
vcf <- read.vcfR("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/Ppin_FRM_baseline_polymorphic_correct_REF_ALT.vcf", verbose = FALSE,convertNA=T)
```

**HTML users should refer to the script version to obtain the associated codes.** 
```{r Define ref and alt alleles VCF, eval=FALSE, include=FALSE}
vcf_fixed <- as.data.frame(vcf@fix)

#Keep only SNP ID, REF, and ALT
snp_info <- vcf_fixed %>%
  dplyr::select(ID, REF, ALT)

data_Lacanau_tot_test <- data_Lacanau_tot
data_Tocchi_tot_test <- data_Tocchi_tot

data_Lacanau_tot_test$ID <- rownames(data_Lacanau_tot_test)
data_Tocchi_tot_test$ID <- rownames(data_Tocchi_tot_test)

convert_genotypes <- function(df, snp_info) {
  df_long <- df %>%
  pivot_longer(-POP, names_to = "ID", values_to = "Genotype")

  #Join with SNP info
  df_long <- df_long %>%
    left_join(snp_info, by = "ID")

  #Create function to classify genotype
  classify <- function(geno, ref, alt) {
    if (is.na(geno) || is.na(ref) || is.na(alt)) return(NA)
    alleles <- strsplit(geno, "")[[1]]
    if (length(alleles) != 2) return(NA)

    if (all(alleles == ref)) return("0/0")
    else if (all(alleles == alt)) return("1/1")
    else if (ref %in% alleles && alt %in% alleles) return("0/1")
    else return(NA)
  }

  df_long <- df_long %>%
    rowwise() %>%
    dplyr::mutate(Genotype_numeric = classify(Genotype, REF, ALT)) %>%
    ungroup()

  #Reshape back to classical format
  df_wide <- df_long %>%
    dplyr::select(ID, POP, Genotype_numeric) %>%
    pivot_wider(names_from = POP, values_from = Genotype_numeric)

  #Set SNP ID as rownames
  df_result <- as.data.frame(df_wide)
  rownames(df_result) <- df_result$ID
  df_result$ID <- NULL
df_result <- data.frame(t(df_result))
  return(df_result)
}

df_LAC_nucl_info <- convert_genotypes(data_Lacanau_tot_test, snp_info)

df_TOC_nucl_info <- convert_genotypes(data_Tocchi_tot_test, snp_info)

#There is 293 snps with NA because the SNPs  were not in the VCF so we need to remove them 
sum(sapply(df_LAC_nucl_info, function(x) all(is.na(x))))
sum(sapply(df_TOC_nucl_info, function(x) all(is.na(x))))

#Remove NA
lac_all_na <- sapply(df_LAC_nucl_info, function(x) all(is.na(x)))

#Remove
df_LAC_nucl_info <- df_LAC_nucl_info[, !lac_all_na]

toc_all_na <- sapply(df_TOC_nucl_info, function(x) all(is.na(x)))

df_TOC_nucl_info <- df_TOC_nucl_info[, !toc_all_na]
```

```{r compute allelic frequencies before imputation, eval=FALSE, include=FALSE}
df_Lacanau_good_format <- df_LAC_nucl_info %>% 
  mutate_all(funs(str_replace(., "0/0", "0"))) %>% 
  mutate_all(funs(str_replace(., "0/1", "1"))) %>% 
  mutate_all(funs(str_replace(., "1/1", "2")))

df_Tocchi_good_format <- df_TOC_nucl_info %>% 
  mutate_all(funs(str_replace(., "0/0", "0"))) %>% 
  mutate_all(funs(str_replace(., "0/1", "1"))) %>% 
  mutate_all(funs(str_replace(., "1/1", "2")))
```

Missing data is imputed using the most common genotype within the population, regardless of cohort, to avoid artificially inflating differences between cohorts:
```{r imputation, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
df_Lacanau_numeric <- df_Lacanau_good_format %>%
  dplyr::mutate(across(everything(), ~ as.numeric(.)))

df_Tocchi_numeric <- df_Tocchi_good_format %>%
  dplyr::mutate(across(everything(), ~ as.numeric(.)))

impute_most_common <- function(df) {
  df_imputed <- as.data.frame(lapply(df, function(col) {
    if (is.numeric(col)) {
      most_common <- as.numeric(names(which.max(table(col, useNA = "no"))))
      replace(col, is.na(col), most_common)
    } else {
      col#Skip non-numeric
    }
  }))
  rownames(df_imputed) <- rownames(df)
  return(df_imputed)
}

df_Lacanau_imputed <- impute_most_common(df_Lacanau_numeric)
df_Tocchi_imputed  <- impute_most_common(df_Tocchi_numeric)

mean(is.na(df_Lacanau_numeric))#Proportion missing
mean(is.na(df_Tocchi_numeric))

save(df_Lacanau_numeric,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Lacanau_numeric.Rdata")

save(df_Tocchi_numeric,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Tocchi_numeric.Rdata")

#Final proportions
mean(is.na(df_Lacanau_imputed))#Should be 0
mean(is.na(df_Tocchi_imputed))
```

```{r compute allelic frequencies after imputation, eval=FALSE, include=FALSE}
df_Lacanau_allelic_frequencies <- df_Lacanau_imputed %>% 
  mutate_all(funs(str_replace(., "0", "0"))) %>% 
  mutate_all(funs(str_replace(., "1", "0.5"))) %>% 
  mutate_all(funs(str_replace(., "2", "1")))

df_Lacanau_allelic_frequencies <- df_Lacanau_allelic_frequencies %>%
  dplyr::mutate(across(everything(), ~ as.numeric(.)))

df_Tocchi_allelic_frequencies <- df_Tocchi_imputed %>% 
  mutate_all(funs(str_replace(., "0", "0"))) %>% 
  mutate_all(funs(str_replace(., "1", "0.5"))) %>% 
  mutate_all(funs(str_replace(., "2", "1")))

df_Tocchi_allelic_frequencies <- df_Tocchi_allelic_frequencies %>%
  dplyr::mutate(across(everything(), ~ as.numeric(.)))
```

```{r save the df, eval=FALSE, include=FALSE}
save(df_Lacanau_imputed,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Lacanau_imputed.Rdata")

save(df_Tocchi_imputed,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Tocchi_imputed.Rdata")

#Allelic frequency
save(df_Lacanau_allelic_frequencies,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Lacanau_allelic_frequencies.Rdata")

save(df_Tocchi_allelic_frequencies,file="Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Tocchi_allelic_frequencies.Rdata")
```

```{r load allelic frequencies}
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Lacanau_allelic_frequencies.Rdata")

load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Tocchi_allelic_frequencies.Rdata")
```

Keep only the climate-associated SNPs for each dataset:
```{r one df per SNP set}
list_plots <- c("Lacanau","Tocchi")
list_SNP <- c("outliers_set_LC_common_final","outliers_set_MC_common_final","all_outliers_set_LC_common_unique_final","all_outliers_set_MC_common_unique_final")
list_name_snp <- c("LC","MC","all_outliers_LC","all_outliers_MC")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  data <- get(paste0("df_",plot,"_allelic_frequencies"))
  
  for(i in 1: length(list_SNP)){
    
    snp_set <- list_SNP[i]
    name_set <- list_name_snp[i]
    snp_set <- get(snp_set)
  
    common_snps <- intersect(colnames(data[,-1]), snp_set)
    filtered_data <- data[, common_snps, drop = FALSE]
    
    #Assign the filtered dataframe to a variable with a descriptive name
    assign(paste0("data_", plot, "_", name_set), filtered_data)
  }
}
```

# Compute individual scores for each Lacanau and Tocchi individuals

In this step, we compute the RDA scores for all sampled individuals in the two populations (Lacanau and Tocchi), using their allelic frequencies for climate-associated SNPs. This allows us to compare them in the RDA model with the optimum values for different time periods. 
Here, we use the predict function with the "wa" type and no scaling, matching the way the RDA values are obtained in the RDA model with the 82 range-wide populations (i.e. historical, past, present, and future optima).

```{r RDA scores LAC and TOC individuals}
list_plots <- c("Lacanau","Tocchi")
list_SNP <- c("outliers_set_LC_common_final", "outliers_set_MC_common_final","all_outliers_set_LC_common_unique_final","all_outliers_set_MC_common_unique_final")
list_name_snp <- c("LC","MC","all_outliers_LC","all_outliers_MC")
list_rda <- c("RDA_outliers_LC","RDA_outliers_MC","RDA_all_outliers_LC",'RDA_all_outliers_MC')

for(x in 1: length(list_plots)){
  
  plot <- list_plots[x]
genomic_data <- get(paste0("df_",plot,"_allelic_frequencies"))

for(i in 1: length(list_rda)){
  
  rda_model <- get(list_rda[i])
  set <- get(list_SNP[i])
  name_snp <- list_name_snp[i]
  
 projected_scores <- predict(rda_model, newdata = genomic_data[set], type = "wa", rank = K, scaling = "none")#"none"

  assign(paste0("adaptive_values_",plot,"_",name_snp),projected_scores)
  }
}

write.csv(adaptive_values_Tocchi_all_outliers_MC, file="Results/Chapter3/Pinaster/Comparison_juveniles_adults/numbers/adaptive_values_Tocchi_all_outliers_MC.csv")
```

Use the Kinship test to identify the few juveniles in Tocchi who are closely related to the future optima. 
```{r draft kinship, eval=FALSE, include=FALSE}
kinshipmatrix <- read.csv(
  "Data/Chapter3/Pinus_pinaster/Simulation/Kinship_matrix_Tocchi.csv",
  header = TRUE, sep = ";", dec = ",",
  skip = 2,
  row.names = 1
)

#Check format
head(kinshipmatrix)

# Numerical matrix
kinshipmatrix_num <- as.matrix(kinshipmatrix)
kinshipmatrix_num <- apply(kinshipmatrix_num, 2, as.numeric)

#Check dimensions of the matrix
dim(kinshipmatrix_num)

high_kinship <- kinshipmatrix[kinshipmatrix > 0.3]

#Display the matrix
high_kinship
```

# Plot the results

We can analyse the results to compare the values of individuals within and between populations, as well as comparing them to the values of the other 82 populations across the range. We can also search for differences between cohorts.  

Prepare the plottings: 
```{r loop plotting}
for(x in 1: length(list_plots)){
  
  plot <- list_plots[x]

for(i in 1: length(list_rda)){

  rda_model <- get(list_rda[i])
  set <- get(list_SNP[i])
  name_snp <- list_name_snp[i]
  
  #Adaptive value plots
  #adaptive_values_plot_LIG_origin <- get(paste0("adaptive_values_",name_snp,"_LIG_origin"))
  #adaptive_values_plot_Mid_Holo_6000_origin <- get(paste0("adaptive_values_",name_snp,"_Mid_Holo_6000_origin"))
  
  #adaptive_values_plot_LIG <- get(paste0("adaptive_values_",name_snp,"_LIG"))
  adaptive_values_plot_Mid_Holo_6000 <- get(paste0("adaptive_values_",name_snp,"_Mid_Holo_6000_mean"))
  adaptive_values_plot_past <- get(paste0("adaptive_values_",name_snp,"_past"))
  adaptive_values_plot_present <- get(paste0("adaptive_values_",name_snp,"_present"))
  adaptive_values_plot_future <- get(paste0("adaptive_values_",name_snp,"_future_mean"))
  
  #Adaptive value individuals
  adaptive_value_obs <- get(paste0("adaptive_values_",plot,"_",name_snp))
  }
}
```

Main figure, we flipped RDA2 all outlier MC to match the direction of the others but the information is unchanged. 
```{r good figures, include=FALSE}
#Flip RDA2 all_outlier MC
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP <- c("outliers_set_LC_common_final","all_outliers_set_LC_common_unique_final", "outliers_set_MC_common_final","all_outliers_set_MC_common_unique_final")
list_name_snp <- c("LC","all_outliers_LC","MC","all_outliers_MC")
list_rda <- c("RDA_outliers_LC","RDA_all_outliers_LC","RDA_outliers_MC","RDA_all_outliers_MC")

#Define color mapping for Population + Age
group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red"
)

t_test_summary <- list()

#Precompute max density across all datasets for uniform y-axis
all_density_values <- c()

for (plot in list_plots) {
  for (i in 1:length(list_rda)) {

    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"

    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign Pop_Age
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Present & future adaptive values
    adaptive_values_past<- get(paste0("adaptive_values_", list_name_snp[i], "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", list_name_snp[i], "_present"))
    adaptive_values_future  <- get(paste0("adaptive_values_", list_name_snp[i], "_future_mean"))
    
    adaptive_value_pres_fut <- rbind(
      adaptive_values_past,
      adaptive_values_present,
      adaptive_values_future
    )

    adaptive_value_pres_fut$Pop_Age <- "Optimal value"

    #Flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    
    all_density_values <- c(all_density_values,
                            density(density_df$RDA1)$y,
                            density(density_df$RDA2)$y)
  }
}

max_density <- max(all_density_values) * 1.05

#Main loop
for (plot in list_plots) {
  
  for (i in 1:length(list_rda)) {
    
    rda_model <- get(list_rda[i])
    set <- get(list_SNP[i])
    name_snp <- list_name_snp[i]
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), digits = 1)
    
    #Climatic arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- env_df$RDA2 * arrow_multiplier
    #flip
    env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)
    
    #Adaptive values
    adaptive_values_past<- get(paste0("adaptive_values_", name_snp, "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
    adaptive_values_future  <- get(paste0("adaptive_values_", name_snp, "_future_mean"))
    
    ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)
    past_plot <- adaptive_values_past[adaptive_values_past$Population %in% ids_plot, ]
    present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
    future_plot  <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]
    
    past_plot$Population <- paste0(past_plot$Population, "_past")
    present_plot$Population <- paste0(present_plot$Population, "_present")
    future_plot$Population  <- paste0(future_plot$Population, "_future")
    adaptive_value_pres_fut <- rbind(past_plot,present_plot, future_plot)
    adaptive_value_pres_fut$Group <- "Optimal value"
    adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    adaptive_value_pres_fut$Time <- ifelse(
      grepl("_past$", adaptive_value_pres_fut$Population), "past",
      ifelse(
        grepl("_present$", adaptive_value_pres_fut$Population), "present",
        "future"
      )
    )
#Flip
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign group: Adult vs Juvenile
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)

    #Flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
    
    #RDA plot
    combined_rda_data <- rbind(
      adaptive_value_obs[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% dplyr::rename(ID = Population)
    )
    
    rda1_limits <- range(combined_rda_data$RDA1, na.rm = TRUE)
    rda2_limits <- range(combined_rda_data$RDA2, na.rm = TRUE)
    
    #RDA scatter plot
    rda_plot <- ggplot(combined_rda_data, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
      geom_point(alpha = 0.7, size = 2) +
      geom_text(data = combined_rda_data[combined_rda_data$Pop_Age == "Optimal value", ],
                aes(label = adaptive_value_pres_fut$Time),
                color = "black", size = 3, vjust = -0.8) +
      geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                   arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
      geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
                color = "black", size = 3) +
      scale_color_manual(values = group_colors) +
      theme_minimal() +
      labs(title = paste("RDA space:", plot, name_snp, "SNPs"),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"),
           y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    #print(rda_plot)
    
     pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/RDA_space/RDA_space_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(rda_plot)
    dev.off()
    
    
    built_rda <- ggplot_build(rda_plot)

    rda1_limits <- built_rda$layout$panel_params[[1]]$x.range
    rda2_limits <- built_rda$layout$panel_params[[1]]$y.range
    
    #Density plots
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df_no_opt <- density_df[density_df$Pop_Age != "Optimal value", ]

    opt_RDA1_past <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="past"]
    opt_RDA1_present <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="present"]
    opt_RDA1_future  <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="future"]
    opt_RDA2_past <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="past"]
    opt_RDA2_present <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="present"]
    opt_RDA2_future  <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="future"]

    #RDA1 density plot
    density_plot_RDA1 <- ggplot(density_df_no_opt, aes(x = RDA1, fill = Pop_Age)) +
      geom_density(color = "black", alpha = 0.8, size = 1, cut = 10) +
      geom_vline(xintercept = opt_RDA1_past, linetype = "dotted", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA1_present, linetype = "dashed", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA1_future, linetype = "solid", color = "red", size = 1) +
      scale_fill_manual(values = group_colors) +
      coord_cartesian(xlim = rda1_limits, ylim = c(0, max_density+5))+
      labs(title = paste0("Density of RDA1 ", name_snp, " SNPs ", plot),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"), y = "Density") +
      theme_minimal() +
      theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

    #print(density_plot_RDA1)
    pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/Density_plot/Density_plot_good_extend_reduce_plot_RDA1_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(density_plot_RDA1)
    dev.off()

    #RDA2 density plot
    density_plot_RDA2 <- ggplot(density_df_no_opt, aes(x = RDA2, fill = Pop_Age)) +
      geom_density(color = "black", alpha = 0.8, size = 1, cut = 10) +
      geom_vline(xintercept = opt_RDA2_past, linetype = "dotted", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA2_present, linetype = "dashed", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA2_future, linetype = "solid", color = "red", size = 1) +
      scale_fill_manual(values = group_colors) +
      coord_cartesian(xlim = rda2_limits, ylim = c(0, max_density + 5))+  
      labs(title = paste0("Density of RDA2 ", name_snp, " SNPs ", plot),
           x = paste0("RDA2 (", explained_variance_round[2], "%)"), y = "Density") +
      theme_minimal() +
      theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

    #print(density_plot_RDA2)
    pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/Density_plot/Density_plot_good_extend_reduce_plot_RDA2_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(density_plot_RDA2)
    dev.off()
  }
}
```

```{r Rmarkdown HTML good figures}
#Flip RDA2 all_outlier MC
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP_Rmarkdown <- c("all_outliers_set_MC_common_unique_final")
list_name_snp_Rmarkdown <- c("all_outliers_MC")
list_rda_Rmarkdown <- c("RDA_all_outliers_MC")

#Define color mapping for Population + Age
group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red"
)

t_test_summary <- list()

#Precompute max density across all datasets for uniform y-axis
all_density_values <- c()

for (plot in list_plots) {
  for (i in 1:length(list_rda_Rmarkdown)) {

    flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"

    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp_Rmarkdown[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign Pop_Age
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Present & future adaptive values
    adaptive_values_past<- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_present"))
    adaptive_values_future  <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_future_mean"))
    
    adaptive_value_pres_fut <- rbind(
      adaptive_values_past,
      adaptive_values_present,
      adaptive_values_future
    )

    adaptive_value_pres_fut$Pop_Age <- "Optimal value"

    #Flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    
    all_density_values <- c(all_density_values,
                            density(density_df$RDA1)$y,
                            density(density_df$RDA2)$y)
  }
}

max_density <- max(all_density_values) * 1.05

#Main loop
for (plot in list_plots) {
  
  for (i in 1:length(list_rda_Rmarkdown)) {
    
    rda_model <- get(list_rda_Rmarkdown[i])
    set <- get(list_SNP_Rmarkdown[i])
    name_snp <- list_name_snp_Rmarkdown[i]
    flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), digits = 1)
    
    #Climatic arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- env_df$RDA2 * arrow_multiplier
    #flip
    env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)
    
    #Adaptive values
    adaptive_values_past<- get(paste0("adaptive_values_", name_snp, "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
    adaptive_values_future  <- get(paste0("adaptive_values_", name_snp, "_future_mean"))
    
    ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)
    past_plot <- adaptive_values_past[adaptive_values_past$Population %in% ids_plot, ]
    present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
    future_plot  <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]
    
    past_plot$Population <- paste0(past_plot$Population, "_past")
    present_plot$Population <- paste0(present_plot$Population, "_present")
    future_plot$Population  <- paste0(future_plot$Population, "_future")
    adaptive_value_pres_fut <- rbind(past_plot,present_plot, future_plot)
    adaptive_value_pres_fut$Group <- "Optimal value"
    adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    adaptive_value_pres_fut$Time <- ifelse(
      grepl("_past$", adaptive_value_pres_fut$Population), "past",
      ifelse(
        grepl("_present$", adaptive_value_pres_fut$Population), "present",
        "future"
      )
    )
#Flip
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign group: Adult vs Juvenile
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)

    #Flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
    
    #RDA plot
    combined_rda_data <- rbind(
      adaptive_value_obs[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% dplyr::rename(ID = Population)
    )
    
    rda1_limits <- range(combined_rda_data$RDA1, na.rm = TRUE)
    rda2_limits <- range(combined_rda_data$RDA2, na.rm = TRUE)
    
    #RDA scatter plot
    rda_plot <- ggplot(combined_rda_data, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
      geom_point(alpha = 0.7, size = 2) +
      geom_text(data = combined_rda_data[combined_rda_data$Pop_Age == "Optimal value", ],
                aes(label = adaptive_value_pres_fut$Time),
                color = "black", size = 3, vjust = -0.8) +
      geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                   arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
      geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
                color = "black", size = 3) +
      scale_color_manual(values = group_colors) +
      theme_minimal() +
      labs(title = paste("RDA space:", plot, name_snp, "SNPs"),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"),
           y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    print(rda_plot)
    
    built_rda <- ggplot_build(rda_plot)

    rda1_limits <- built_rda$layout$panel_params[[1]]$x.range
    rda2_limits <- built_rda$layout$panel_params[[1]]$y.range
    
    #Density plots
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df_no_opt <- density_df[density_df$Pop_Age != "Optimal value", ]

    opt_RDA1_past <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="past"]
    opt_RDA1_present <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="present"]
    opt_RDA1_future  <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="future"]
    opt_RDA2_past <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="past"]
    opt_RDA2_present <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="present"]
    opt_RDA2_future  <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="future"]

    #RDA1 density plot
    density_plot_RDA1 <- ggplot(density_df_no_opt, aes(x = RDA1, fill = Pop_Age)) +
      geom_density(color = "black", alpha = 0.8, size = 1, cut = 10) +
      geom_vline(xintercept = opt_RDA1_past, linetype = "dotted", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA1_present, linetype = "dashed", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA1_future, linetype = "solid", color = "red", size = 1) +
      scale_fill_manual(values = group_colors) +
      coord_cartesian(xlim = rda1_limits, ylim = c(0, max_density+5))+
      labs(title = paste0("Density of RDA1 ", name_snp, " SNPs ", plot),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"), y = "Density") +
      theme_minimal() +
      theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

    print(density_plot_RDA1)

    #RDA2 density plot
    density_plot_RDA2 <- ggplot(density_df_no_opt, aes(x = RDA2, fill = Pop_Age)) +
      geom_density(color = "black", alpha = 0.8, size = 1, cut = 10) +
      geom_vline(xintercept = opt_RDA2_past, linetype = "dotted", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA2_present, linetype = "dashed", color = "red", size = 1) +
      geom_vline(xintercept = opt_RDA2_future, linetype = "solid", color = "red", size = 1) +
      scale_fill_manual(values = group_colors) +
      coord_cartesian(xlim = rda2_limits, ylim = c(0, max_density + 5))+  
      labs(title = paste0("Density of RDA2 ", name_snp, " SNPs ", plot),
           x = paste0("RDA2 (", explained_variance_round[2], "%)"), y = "Density") +
      theme_minimal() +
      theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

    print(density_plot_RDA2)
  }
}
```

Check if genetic shifts between adults and juveniles are significants using a t-test: 
```{r statistics}
t_test_summary <- list()

flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

for (plot in list_plots) {
  
  for (i in 1:length(list_rda)) {
    
    rda_model <- get(list_rda[i])
    set <- get(list_SNP[i])
    name_snp <- list_name_snp[i]
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), digits = 1)
    
    #Climatic arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- env_df$RDA2 * arrow_multiplier
    env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)
    
    #Adaptive values
    adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
    adaptive_values_future <- get(paste0("adaptive_values_", name_snp, "_future_mean"))
    
    ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)
    present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
    future_plot <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]
    
    #Merge present & future
    present_plot$Population <- paste0(present_plot$Population, "_present")
    future_plot$Population <- paste0(future_plot$Population, "_future")
    adaptive_value_pres_fut <- rbind(present_plot, future_plot)
    adaptive_value_pres_fut$Group <- "Optimal value"
    adaptive_value_pres_fut$Pop_Age <- "Optimal value"  # for consistent rbind
    adaptive_value_pres_fut$Time <- ifelse(grepl("_present$", adaptive_value_pres_fut$Population), "present", "future")
    
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign group: Adult vs Juvenile
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
    
    #RDA plots
    combined_rda_data <- rbind(
      adaptive_value_obs[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% 
        dplyr::rename(ID = Population)
    )
    
    #Combine for density: only Adults & Juveniles
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    
    #T-test
    df_ttest <- density_df
    for (axis in c("RDA1", "RDA2")) {
      ttest <- t.test(as.formula(paste(axis, "~ Pop_Age")), data = df_ttest)
      t_test_summary[[length(t_test_summary)+1]] <- data.frame(
        Plot = plot,
        SNP_Set = name_snp,
        Axis = axis,
        t_statistic = ttest$statistic,
        p_value = ttest$p.value,
        mean_juvenile = mean(df_ttest[[axis]][df_ttest$Pop_Age %in% paste0(plot,"_Juvenile")], na.rm = TRUE),
        mean_adult = mean(df_ttest[[axis]][df_ttest$Pop_Age %in% paste0(plot,"_Adult")], na.rm = TRUE)
      )
    }
    
  }
}

#Combine t-test results
t_test_results_df <- do.call(rbind, t_test_summary)
print(t_test_results_df)
```

Include LIG and Mid-holocene climates:
```{r good figures Mid-holocene, include=FALSE}
#Flip RDA2 all_outlier MC
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP <- c("outliers_set_LC_common_final","all_outliers_set_LC_common_unique_final", "outliers_set_MC_common_final","all_outliers_set_MC_common_unique_final")
list_name_snp <- c("LC","all_outliers_LC","MC","all_outliers_MC")
list_rda <- c("RDA_outliers_LC","RDA_all_outliers_LC","RDA_outliers_MC","RDA_all_outliers_MC")

#Define color mapping for Population + Age
group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red"
)

#Precompute max density across all datasets for uniform y-axis
all_density_values <- c()

for (plot in list_plots) {
  for (i in 1:length(list_rda)) {
    
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"

    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign Pop_Age
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Present & future adaptive values
#adaptive_values_LIG  <- get(paste0("adaptive_values_", list_name_snp[i], "_LIG"))
adaptive_values_Mid_Holo_6000 <- get(paste0("adaptive_values_", list_name_snp[i], "_Mid_Holo_6000_mean"))
adaptive_values_past <- get(paste0("adaptive_values_", list_name_snp[i], "_past"))
adaptive_values_present <- get(paste0("adaptive_values_", list_name_snp[i], "_present"))
adaptive_values_future  <- get(paste0("adaptive_values_", list_name_snp[i], "_future_mean"))

#Add Time column before binding
#adaptive_values_LIG$Time  <- "LIG"
adaptive_values_Mid_Holo_6000$Time <- "Mid_Holo_6000"
adaptive_values_past$Time <- "past"
adaptive_values_present$Time <- "present"
adaptive_values_future$Time <- "future"

#Bind everything
adaptive_value_pres_fut <- rbind(
  #adaptive_values_LIG,
  adaptive_values_Mid_Holo_6000,
  adaptive_values_past,
  adaptive_values_present,
  adaptive_values_future
)

#Add Pop_Age column
adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    
adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Combine for density
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    
    #Compute density values
    all_density_values <- c(all_density_values,
                            density(density_df$RDA1)$y,
                            density(density_df$RDA2)$y)
  }
}

#Maximum density for y-axis
max_density <- max(all_density_values) * 1.05

#Main loop
for (plot in list_plots) {
  
  for (i in 1:length(list_rda)) {
    
    rda_model <- get(list_rda[i])
    set <- get(list_SNP[i])
    name_snp <- list_name_snp[i]
     flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), digits = 1)
    
    #Climatic arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- env_df$RDA2 * arrow_multiplier
    
    env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)

    #Adaptive values
 #adaptive_values_LIG  <- get(paste0("adaptive_values_", name_snp, "_LIG"))
adaptive_values_Mid_Holo_6000 <- get(paste0("adaptive_values_", name_snp, "_Mid_Holo_6000_mean"))
adaptive_values_past <- get(paste0("adaptive_values_", name_snp, "_past"))
adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
adaptive_values_future  <- get(paste0("adaptive_values_", name_snp, "_future_mean"))

ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)

#LIG_plot  <- adaptive_values_LIG[adaptive_values_LIG$Population %in% ids_plot, ]
Mid_Holo_6000_plot <- adaptive_values_Mid_Holo_6000[adaptive_values_Mid_Holo_6000$Population %in% ids_plot, ]
past_plot <- adaptive_values_past[adaptive_values_past$Population %in% ids_plot, ]
present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
future_plot  <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]

#Assign Time explicitly (no regex needed)
#LIG_plot$Time  <- "LIG"
Mid_Holo_6000_plot$Time <- "Mid_Holo_6000"
past_plot$Time <- "past"
present_plot$Time <- "present"
future_plot$Time <- "future"

#LIG_plot$Population  <- paste0(LIG_plot$Population, "_LIG")
Mid_Holo_6000_plot$Population <- paste0(Mid_Holo_6000_plot$Population, "_Mid_Holo_6000")
past_plot$Population <- paste0(past_plot$Population, "_past")
present_plot$Population <- paste0(present_plot$Population, "_present")
future_plot$Population  <- paste0(future_plot$Population, "_future")

adaptive_value_pres_fut <- rbind(
  #LIG_plot,
  Mid_Holo_6000_plot,
  past_plot,
  present_plot,
  future_plot
)

adaptive_value_pres_fut$Group   <- "Optimal value"
adaptive_value_pres_fut$Pop_Age <- "Optimal value"

#Flip
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign group: Adult vs Juvenile
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    #flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)

    
    #RDA plot
    combined_rda_data <- rbind(
      adaptive_value_obs[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% dplyr::rename(ID = Population)
    )
    
    rda1_limits <- range(combined_rda_data$RDA1, na.rm = TRUE)
    rda2_limits <- range(combined_rda_data$RDA2, na.rm = TRUE)
    
    #RDA scatter plot
    rda_plot <- ggplot(combined_rda_data, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
      geom_point(alpha = 0.7, size = 3) +
      geom_text(data = combined_rda_data[combined_rda_data$Pop_Age == "Optimal value", ],
                aes(label = adaptive_value_pres_fut$Time),
                color = "black", size = 2, vjust = -0.8) +#3
      geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                   arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
      geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
                color = "black", size = 3) +
      scale_color_manual(values = group_colors) +
      theme_minimal() +
      labs(title = paste("RDA space:", plot, name_snp, "SNPs"),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"),
           y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  #  print(rda_plot)
    
     pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/Historical_climates/RDA_space_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(rda_plot)
    dev.off()
    
    built_rda <- ggplot_build(rda_plot)

rda1_limits <- built_rda$layout$panel_params[[1]]$x.range
rda2_limits <- built_rda$layout$panel_params[[1]]$y.range
    
    #Density plots
    density_df <- rbind(
  adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
  adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
)
density_df_no_opt <- density_df[density_df$Pop_Age != "Optimal value", ]

#Optimal lines
#opt_RDA1_LIG <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="LIG"]
opt_RDA1_Mid_Holo_6000 <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="Mid_Holo_6000"]
opt_RDA1_past <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="past"]
opt_RDA1_present <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="present"]
opt_RDA1_future  <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="future"]

#opt_RDA2_LIG <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="LIG"]
opt_RDA2_Mid_Holo_6000 <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="Mid_Holo_6000"]
opt_RDA2_past <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="past"]
opt_RDA2_present <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="present"]
opt_RDA2_future  <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="future"]

#RDA1 density plot
density_plot_RDA1 <- ggplot(density_df_no_opt, aes(x = RDA1, fill = Pop_Age)) +
  geom_density(
  color = "black",
  alpha = 0.8,
  size = 1,
  cut = 10
) +
#  geom_vline(xintercept = opt_RDA1_LIG, linetype = "solid", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA1_Mid_Holo_6000, linetype = "dotted", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA1_past, linetype = "dotted", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA1_present, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA1_future, linetype = "solid", color = "red", size = 1) +
  scale_fill_manual(values = group_colors) +
  coord_cartesian(
  xlim = rda1_limits,
  ylim = c(0, max_density+5)
)+
  labs(title = paste0("Density of RDA1 ", name_snp, " SNPs ", plot),
       x = paste0("RDA1 (", explained_variance_round[1], "%)"), y = "Density") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

#print(density_plot_RDA1)
pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/Historical_climates/Density_plot_good_extend_reduce_plot_RDA1_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(density_plot_RDA1)
    dev.off()

#RDA2 density plot
density_plot_RDA2 <- ggplot(density_df_no_opt, aes(x = RDA2, fill = Pop_Age)) +
  geom_density(
  color = "black",
  alpha = 0.8,
  size = 1,
  cut = 10
) +
#  geom_vline(xintercept = opt_RDA2_LIG, linetype = "solid", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA2_Mid_Holo_6000, linetype = "dotted", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA2_past, linetype = "dotted", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA2_present, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA2_future, linetype = "solid", color = "red", size = 1) +
  scale_fill_manual(values = group_colors) +
 coord_cartesian(
  xlim = rda2_limits,
  ylim = c(0, max_density + 5)
)+  
  labs(title = paste0("Density of RDA2 ", name_snp, " SNPs ", plot),
       x = paste0("RDA2 (", explained_variance_round[2], "%)"), y = "Density") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

#print(density_plot_RDA2)
pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/Historical_climates/Density_plot_good_extend_reduce_plot_RDA2_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(density_plot_RDA2)
    dev.off()
    }
  }
```

```{r Rmarkdown good figures Mid-holocene}
#Flip RDA2 all_outlier MC
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP_Rmarkdown <- c("all_outliers_set_MC_common_unique_final")
list_name_snp_Rmarkdown <- c("all_outliers_MC")
list_rda_Rmarkdown <- c("RDA_all_outliers_MC")

#Define color mapping for Population + Age
group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red"
)

#Precompute max density across all datasets for uniform y-axis
all_density_values <- c()

for (plot in list_plots) {
  for (i in 1:length(list_rda_Rmarkdown)) {
    
    flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"

    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp_Rmarkdown[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign Pop_Age
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Present & future adaptive values
#adaptive_values_LIG  <- get(paste0("adaptive_values_", list_name_snp[i], "_LIG"))
adaptive_values_Mid_Holo_6000 <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_Mid_Holo_6000_mean"))
adaptive_values_past <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_past"))
adaptive_values_present <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_present"))
adaptive_values_future  <- get(paste0("adaptive_values_", list_name_snp_Rmarkdown[i], "_future_mean"))

#Add Time column before binding
#adaptive_values_LIG$Time  <- "LIG"
adaptive_values_Mid_Holo_6000$Time <- "Mid_Holo_6000"
adaptive_values_past$Time <- "past"
adaptive_values_present$Time <- "present"
adaptive_values_future$Time <- "future"

#Bind everything
adaptive_value_pres_fut <- rbind(
  #adaptive_values_LIG,
  adaptive_values_Mid_Holo_6000,
  adaptive_values_past,
  adaptive_values_present,
  adaptive_values_future
)

#Add Pop_Age column
adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    
adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)
adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Combine for density
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    
    #Compute density values
    all_density_values <- c(all_density_values,
                            density(density_df$RDA1)$y,
                            density(density_df$RDA2)$y)
  }
}

#Maximum density for y-axis
max_density <- max(all_density_values) * 1.05

#Main loop
for (plot in list_plots) {
  
  for (i in 1:length(list_rda_Rmarkdown)) {
    
    rda_model <- get(list_rda_Rmarkdown[i])
    set <- get(list_SNP_Rmarkdown[i])
    name_snp <- list_name_snp_Rmarkdown[i]
     flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), digits = 1)
    
    #Climatic arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- env_df$RDA2 * arrow_multiplier
    
    env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)

    #Adaptive values
 #adaptive_values_LIG  <- get(paste0("adaptive_values_", name_snp, "_LIG"))
adaptive_values_Mid_Holo_6000 <- get(paste0("adaptive_values_", name_snp, "_Mid_Holo_6000_mean"))
adaptive_values_past <- get(paste0("adaptive_values_", name_snp, "_past"))
adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
adaptive_values_future  <- get(paste0("adaptive_values_", name_snp, "_future_mean"))

ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)

#LIG_plot  <- adaptive_values_LIG[adaptive_values_LIG$Population %in% ids_plot, ]
Mid_Holo_6000_plot <- adaptive_values_Mid_Holo_6000[adaptive_values_Mid_Holo_6000$Population %in% ids_plot, ]
past_plot <- adaptive_values_past[adaptive_values_past$Population %in% ids_plot, ]
present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
future_plot  <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]

#Assign Time explicitly (no regex needed)
#LIG_plot$Time  <- "LIG"
Mid_Holo_6000_plot$Time <- "Mid_Holo_6000"
past_plot$Time <- "past"
present_plot$Time <- "present"
future_plot$Time <- "future"

#LIG_plot$Population  <- paste0(LIG_plot$Population, "_LIG")
Mid_Holo_6000_plot$Population <- paste0(Mid_Holo_6000_plot$Population, "_Mid_Holo_6000")
past_plot$Population <- paste0(past_plot$Population, "_past")
present_plot$Population <- paste0(present_plot$Population, "_present")
future_plot$Population  <- paste0(future_plot$Population, "_future")

adaptive_value_pres_fut <- rbind(
  #LIG_plot,
  Mid_Holo_6000_plot,
  past_plot,
  present_plot,
  future_plot
)

adaptive_value_pres_fut$Group   <- "Optimal value"
adaptive_value_pres_fut$Pop_Age <- "Optimal value"

#Flip
    adaptive_value_pres_fut$RDA2 <- flip_if_needed(adaptive_value_pres_fut$RDA2, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    #Assign group: Adult vs Juvenile
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    #flip
    adaptive_value_obs$RDA2 <- flip_if_needed(adaptive_value_obs$RDA2, flip_RDA2)

    #RDA plot
    combined_rda_data <- rbind(
      adaptive_value_obs[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% dplyr::rename(ID = Population)
    )
    
    rda1_limits <- range(combined_rda_data$RDA1, na.rm = TRUE)
    rda2_limits <- range(combined_rda_data$RDA2, na.rm = TRUE)
    
    #RDA scatter plot
    rda_plot <- ggplot(combined_rda_data, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
      geom_point(alpha = 0.7, size = 3) +
      geom_text(data = combined_rda_data[combined_rda_data$Pop_Age == "Optimal value", ],
                aes(label = adaptive_value_pres_fut$Time),
                color = "black", size = 2, vjust = -0.8) +#3
      geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                   arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
      geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
                color = "black", size = 3) +
      scale_color_manual(values = group_colors) +
      theme_minimal() +
      labs(title = paste("RDA space:", plot, name_snp, "SNPs"),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"),
           y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    print(rda_plot)
    
    built_rda <- ggplot_build(rda_plot)

rda1_limits <- built_rda$layout$panel_params[[1]]$x.range
rda2_limits <- built_rda$layout$panel_params[[1]]$y.range
    
    #Density plots
    density_df <- rbind(
  adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
  adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
)
density_df_no_opt <- density_df[density_df$Pop_Age != "Optimal value", ]

#Optimal lines
#opt_RDA1_LIG <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="LIG"]
opt_RDA1_Mid_Holo_6000 <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="Mid_Holo_6000"]
opt_RDA1_past <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="past"]
opt_RDA1_present <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="present"]
opt_RDA1_future  <- adaptive_value_pres_fut$RDA1[adaptive_value_pres_fut$Time=="future"]

#opt_RDA2_LIG <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="LIG"]
opt_RDA2_Mid_Holo_6000 <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="Mid_Holo_6000"]
opt_RDA2_past <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="past"]
opt_RDA2_present <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="present"]
opt_RDA2_future  <- adaptive_value_pres_fut$RDA2[adaptive_value_pres_fut$Time=="future"]

#RDA1 density plot
density_plot_RDA1 <- ggplot(density_df_no_opt, aes(x = RDA1, fill = Pop_Age)) +
  geom_density(
  color = "black",
  alpha = 0.8,
  size = 1,
  cut = 10
) +
#  geom_vline(xintercept = opt_RDA1_LIG, linetype = "solid", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA1_Mid_Holo_6000, linetype = "dotted", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA1_past, linetype = "dotted", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA1_present, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA1_future, linetype = "solid", color = "red", size = 1) +
  scale_fill_manual(values = group_colors) +
  coord_cartesian(
  xlim = rda1_limits,
  ylim = c(0, max_density+5)
)+
  labs(title = paste0("Density of RDA1 ", name_snp, " SNPs ", plot),
       x = paste0("RDA1 (", explained_variance_round[1], "%)"), y = "Density") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

print(density_plot_RDA1)

#RDA2 density plot
density_plot_RDA2 <- ggplot(density_df_no_opt, aes(x = RDA2, fill = Pop_Age)) +
  geom_density(
  color = "black",
  alpha = 0.8,
  size = 1,
  cut = 10
) +
#  geom_vline(xintercept = opt_RDA2_LIG, linetype = "solid", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA2_Mid_Holo_6000, linetype = "dotted", color = "blue", size = 1) +
  geom_vline(xintercept = opt_RDA2_past, linetype = "dotted", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA2_present, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = opt_RDA2_future, linetype = "solid", color = "red", size = 1) +
  scale_fill_manual(values = group_colors) +
 coord_cartesian(
  xlim = rda2_limits,
  ylim = c(0, max_density + 5)
)+  
  labs(title = paste0("Density of RDA2 ", name_snp, " SNPs ", plot),
       x = paste0("RDA2 (", explained_variance_round[2], "%)"), y = "Density") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.line = element_line(colour = "black"))

print(density_plot_RDA2)
    }
  }
```
Interestingly, the LIG climate is closer to the observed climate than to the climate of the Mid Holocene 6,000 years BP.
The patterns make sense for Lacanau, as it seems that there have been gradual climatic changes in the same direction, except during the Mid Holocene, when the climate was more extreme. 
In Tocchi, it seems that the trees were never adapted to their environment. However, it seems that climate change is reducing the adaptation lag, either by bringing it closer to adaptation, or because the trees were not adapted at the LIG and Mid Holocene due to the climate being too different from their origin, but are now adapting to climate change. 

We can look at the position of the populations used to build the model for comparison:
```{r position pop rda space, include=FALSE}
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_rda <- c("RDA_outliers_LC",
              "RDA_all_outliers_LC",
              "RDA_outliers_MC",
              "RDA_all_outliers_MC")

list_name_snp <- c("LC",
                   "all_outliers_LC",
                   "MC",
                   "all_outliers_MC")

for (i in seq_along(list_rda)) {
  
  rda_model <- get(list_rda[i])
  name_snp  <- list_name_snp[i]
  
  explained_variance <- round(
    rda_model$CCA$eig * 100 / sum(rda_model$CCA$eig), 1
  )
  
  flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
  
  #LC site scores (82 populations)
  lc_scores <- scores(rda_model,
                      display = "lc",
                      scaling = "none",
                      choices = 1:2)
  
  lc_df <- as.data.frame(lc_scores)
  
  #Flip RDA2 if needed
  lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
  
  #Add Population ID and Country
  lc_df$Population_ID <- rownames(lc_df)
  lc_df$Country <- meta_data_pop_order$Country
  
  #Environmental arrows
  env_scores <- scores(rda_model,
                       display = "bp",
                       scaling = "none",
                       choices = 1:2)
  
  env_df <- as.data.frame(env_scores)
  env_df$Variable <- rownames(env_df)
  
  arrow_multiplier <- 0.15
  env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
  env_df$RDA2 <- flip_if_needed(env_df$RDA2 * arrow_multiplier, flip_RDA2)
  
  #Plot
  p <- ggplot(lc_df, aes(x = RDA1, y = RDA2, color = Country)) +
    geom_point(size = 2.5, alpha = 0.9) +
    
    #Optionally add population labels
    geom_text(aes(label = Population_ID),
              size = 3,
              vjust = -0.5,
              check_overlap = TRUE) +
    
    #Environmental arrows
    geom_segment(data = env_df,
                 aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                 arrow = arrow(length = unit(0.25, "cm")),
                 color = "black",
                 inherit.aes = FALSE) +
    
    geom_text(data = env_df,
              aes(x = RDA1 * 0.9,
                  y = RDA2 * 0.9,
                  label = Variable),
              size = 4,
              inherit.aes = FALSE) +
    
    theme_minimal() +
    labs(title = paste("Climatic RDA space –", name_snp),
         x = paste0("RDA1 (", explained_variance[1], "%)"),
         y = paste0("RDA2 (", explained_variance[2], "%)"),
         color = "Country") +
    
    theme(panel.grid = element_blank(),
          axis.line = element_line(colour = "black"),
          plot.title = element_text(hjust = 0.5))
  
 # print(p)
  pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/RDA_space_all_individuals_", name_snp, "_set.pdf"),
      width = 8, height = 6)
  print(p)
  dev.off()
}
```

With Lacanau and Tocchi individuals
```{r good figures with individuals}
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP <- c("outliers_set_LC_common_final",
              "all_outliers_set_LC_common_unique_final", 
              "outliers_set_MC_common_final",
              "all_outliers_set_MC_common_unique_final")

list_name_snp <- c("LC","all_outliers_LC","MC","all_outliers_MC")
list_rda <- c("RDA_outliers_LC","RDA_all_outliers_LC","RDA_outliers_MC","RDA_all_outliers_MC")

#Color mapping
group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red", 
  "ALG"="green",
  "MAR"="darkgreen",
  "ESP"="pink",
  "FRA"="purple",
  "ITA"="brown",
  "PRT"="black",
  "TUN"="red"
)

t_test_summary <- list()
all_density_values <- c()

#Precompute density values for uniform y-axis
for (plot in list_plots) {
  for (i in seq_along(list_rda)) {
    
    rda_model <- get(list_rda[i])
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = flip_if_needed(adaptive_value_obs[,2], flip_RDA2)
    )
    
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Past / Present / Future
    adaptive_values_past <- get(paste0("adaptive_values_", list_name_snp[i], "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", list_name_snp[i], "_present"))
    adaptive_values_future <- get(paste0("adaptive_values_", list_name_snp[i], "_future_mean"))
    
    for (df in list(adaptive_values_past, adaptive_values_present, adaptive_values_future)) {
      df$RDA2 <- flip_if_needed(df$RDA2, flip_RDA2)
    }
    
    adaptive_value_pres_fut <- rbind(adaptive_values_past, adaptive_values_present, adaptive_values_future)
    adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    adaptive_value_pres_fut$Time <- ifelse(
      grepl("_past$", adaptive_value_pres_fut$Population), "past",
      ifelse(grepl("_present$", adaptive_value_pres_fut$Population), "present", "future")
    )
    
    #Combine density values for y-axis scaling
    density_df <- rbind(
      adaptive_value_obs[, c("RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("RDA1","RDA2","Pop_Age")]
    )
    density_df <- density_df[density_df$Pop_Age != "Optimal value", ]
    all_density_values <- c(all_density_values,
                            density(density_df$RDA1)$y,
                            density(density_df$RDA2)$y)
  }
}

max_density <- max(all_density_values) * 1.05

#Main plotting loop
for (plot in list_plots) {
  for (i in seq_along(list_rda)) {
    
    rda_model <- get(list_rda[i])
    set <- get(list_SNP[i])
    name_snp <- list_name_snp[i]
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), 1)
    
    #Environmental arrows
    env_scores <- scores(rda_model, display = "bp")
    env_df <- as.data.frame(env_scores)
    env_df$Variable <- rownames(env_df)
    arrow_multiplier <- 0.15
    env_df$RDA1 <- env_df$RDA1 * arrow_multiplier
    env_df$RDA2 <- flip_if_needed(env_df$RDA2 * arrow_multiplier, flip_RDA2)
    
    #Observed adaptive values
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", name_snp))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = flip_if_needed(adaptive_value_obs[,2], flip_RDA2)
    )
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    #Range-wide LC scores
    lc_scores <- scores(rda_model, display = "lc", scaling = "none", choices = 1:2)
    lc_df <- as.data.frame(lc_scores)
    lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
    lc_df$Population_ID <- rownames(lc_df)
    lc_df$Country <- meta_data_pop_order$Country
    adaptive_values_range_wide_final <- data.frame(
      ID= lc_df$Population_ID,
      RDA1=lc_df$RDA1,
      RDA2=lc_df$RDA2,
      Group="range_wide",
      Pop_Age=lc_df$Country
    )
    
    adaptive_value_obs_final <- rbind(adaptive_value_obs, adaptive_values_range_wide_final)
    
    #Past / Present / Future adaptive values
    adaptive_values_past <- get(paste0("adaptive_values_", name_snp, "_past"))
    adaptive_values_present <- get(paste0("adaptive_values_", name_snp, "_present"))
    adaptive_values_future <- get(paste0("adaptive_values_", name_snp, "_future_mean"))
    
    adaptive_values_past$RDA2    <- flip_if_needed(adaptive_values_past$RDA2, flip_RDA2)
adaptive_values_present$RDA2 <- flip_if_needed(adaptive_values_present$RDA2, flip_RDA2)
adaptive_values_future$RDA2  <- flip_if_needed(adaptive_values_future$RDA2, flip_RDA2)

    ids_plot <- grep(plot, adaptive_values_present$Population, value = TRUE)
    past_plot <- adaptive_values_past[adaptive_values_past$Population %in% ids_plot, ]
    present_plot <- adaptive_values_present[adaptive_values_present$Population %in% ids_plot, ]
    future_plot <- adaptive_values_future[adaptive_values_future$Population %in% ids_plot, ]
    past_plot$Population <- paste0(past_plot$Population, "_past")
    present_plot$Population <- paste0(present_plot$Population, "_present")
    future_plot$Population <- paste0(future_plot$Population, "_future")
    adaptive_value_pres_fut <- rbind(past_plot,present_plot,future_plot)
    adaptive_value_pres_fut$Group <- "Optimal value"
    adaptive_value_pres_fut$Pop_Age <- "Optimal value"
    adaptive_value_pres_fut$Time <- ifelse(
      grepl("_past$", adaptive_value_pres_fut$Population), "past",
      ifelse(grepl("_present$", adaptive_value_pres_fut$Population), "present","future")
    )
    
    #Combine data for RDA scatter
    combined_rda_data <- rbind(
      adaptive_value_obs_final[, c("ID","RDA1","RDA2","Pop_Age")],
      adaptive_value_pres_fut[, c("Population","RDA1","RDA2","Pop_Age")] %>% dplyr::rename(ID = Population)
    )
    
    rda_plot <- ggplot(combined_rda_data, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
      geom_point(alpha = 0.7, size = 2) +
      geom_text(data = combined_rda_data[combined_rda_data$Pop_Age == "Optimal value", ],
                aes(label = adaptive_value_pres_fut$Time),
                color = "black", size = 3, vjust = -0.8) +
      geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                   arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
      geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
                color = "black", size = 3) +
      scale_color_manual(values = group_colors) +
      theme_minimal() +
      labs(title = paste("RDA space:", plot, name_snp, "SNPs"),
           x = paste0("RDA1 (", explained_variance_round[1], "%)"),
           y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    
    #print(rda_plot)
    
    pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/",
               plot, "/all_individuals/RDA_space/RDA_space_all_pop_", name_snp, "_set_", plot, ".pdf"),
        width = 8, height = 6)
    print(rda_plot)
    dev.off()
  }
}

#Combine t-test results
t_test_results_df <- do.call(rbind, t_test_summary)
print(t_test_results_df)
```

With Lacanau and Tocchi individuals in the same plot
```{r good figures combined plots, include=FALSE}
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP <- c("outliers_set_LC_common_final","all_outliers_set_LC_common_unique_final", "outliers_set_MC_common_final","all_outliers_set_MC_common_unique_final")
list_name_snp <- c("LC","all_outliers_LC","MC","all_outliers_MC")
list_rda <- c("RDA_outliers_LC","RDA_all_outliers_LC","RDA_outliers_MC","RDA_all_outliers_MC")

group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red", 
  "ALG"="green",
  "MAR"="darkgreen",
  "ESP"="pink",
  "FRA"="purple",
  "ITA"="brown",
  "PRT"="black",
  "TUN"="red"
)

#Precompute max density for uniform y-axis
all_density_values <- c()
for (plot in list_plots) {
  for (i in seq_along(list_rda)) {
    flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
    
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    assign(paste0("adaptive_value_obs_",plot,"_",list_name_snp[i]), adaptive_value_obs)
  }
}

#Compute density values across all datasets
for (i in seq_along(list_rda)) {
  rda_model <- get(list_rda[i])
  name_snp  <- list_name_snp[i]
  flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
  
  lc_scores <- scores(rda_model, display = "lc", scaling = "none", choices = 1:2)
  lc_df <- as.data.frame(lc_scores)
  lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
  lc_df$Population_ID <- rownames(lc_df)
  lc_df$Country <- meta_data_pop_order$Country
  
  adaptive_values_range_wide_final <- data.frame(
    ID= lc_df$Population_ID,
    RDA1=lc_df$RDA1,
    RDA2=lc_df$RDA2,
    Group="range_wide",
    Pop_Age=lc_df$Country
  )
  
  adaptive_obs_Lacanau <- get(paste0("adaptive_value_obs_Lacanau_",name_snp))
  adaptive_obs_Tocchi <- get(paste0("adaptive_value_obs_Tocchi_",name_snp))
  
  adaptive_value_obs_final <- rbind(adaptive_obs_Lacanau, adaptive_obs_Tocchi, adaptive_values_range_wide_final)
  
  all_density_values <- c(all_density_values,
                          density(adaptive_value_obs_final$RDA1)$y,
                          density(adaptive_value_obs_final$RDA2)$y)
}

max_density <- max(all_density_values) * 1.05

#Main RDA plotting loop
for (i in seq_along(list_rda)) {
  rda_model <- get(list_rda[i])
  name_snp <- list_name_snp[i]
  flip_RDA2 <- list_rda[i] == "RDA_all_outliers_MC"
  
  explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), 1)
  
  #Environmental arrows
  env_scores <- scores(rda_model, display = "bp")
  env_df <- as.data.frame(env_scores)
  env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)
  env_df$RDA1 <- env_df$RDA1 * 0.15
  env_df$RDA2 <- env_df$RDA2 * 0.15
  env_df$Variable <- rownames(env_df)
  
  adaptive_obs_Lacanau <- get(paste0("adaptive_value_obs_Lacanau_",name_snp))
  adaptive_obs_Tocchi <- get(paste0("adaptive_value_obs_Tocchi_",name_snp))
  adaptive_obs_Lacanau$RDA2 <- flip_if_needed(adaptive_obs_Lacanau$RDA2, flip_RDA2)
  adaptive_obs_Tocchi$RDA2 <- flip_if_needed(adaptive_obs_Tocchi$RDA2, flip_RDA2)
  
  lc_scores <- scores(rda_model, display = "lc", scaling = "none", choices = 1:2)
  lc_df <- as.data.frame(lc_scores)
  lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
  lc_df$Population_ID <- rownames(lc_df)
  lc_df$Country <- meta_data_pop_order$Country
  adaptive_values_range_wide_final <- data.frame(
    ID= lc_df$Population_ID,
    RDA1=lc_df$RDA1,
    RDA2=lc_df$RDA2,
    Group="range_wide",
    Pop_Age=lc_df$Country
  )
  
  adaptive_value_obs_final <- rbind(adaptive_obs_Lacanau, adaptive_obs_Tocchi, adaptive_values_range_wide_final)
  
  rda_plot <- ggplot(adaptive_value_obs_final, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                 arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
    geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
              color = "black", size = 3) +
    scale_color_manual(values = group_colors) +
    theme_minimal() +
    labs(title = paste("RDA space:", name_snp, "SNPs"),
         x = paste0("RDA1 (", explained_variance_round[1], "%)"),
         y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
    theme(legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
  #print(rda_plot)
  
  pdf(paste0("Results/Chapter3/Pinaster/Comparison_juveniles_adults/figures/RDA_space_all_pop_LAC_TOC", name_snp, "_set.pdf"),
      width = 8, height = 6)
  print(rda_plot)
  dev.off()
}
```

```{r Rmarkdown good figures combined plots}
flip_if_needed <- function(x, flip) {
  if (flip) return(-x)
  return(x)
}

list_SNP_Rmarkdown <- c("all_outliers_set_MC_common_unique_final")
list_name_snp_Rmarkdown <- c("all_outliers_MC")
list_rda_Rmarkdown <- c("RDA_all_outliers_MC")

group_colors <- c(
  "Lacanau_Adult" = "#3A6FB0",
  "Lacanau_Juvenile" = "#A7C7E7",
  "Tocchi_Adult" = "#C49A00",
  "Tocchi_Juvenile" = "#FFD700",
  "Optimal value" = "red", 
  "ALG"="green",
  "MAR"="darkgreen",
  "ESP"="pink",
  "FRA"="purple",
  "ITA"="brown",
  "PRT"="black",
  "TUN"="red"
)

#Precompute max density for uniform y-axis
all_density_values <- c()
for (plot in list_plots) {
  for (i in seq_along(list_rda_Rmarkdown)) {
    flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"
    
    adaptive_value_obs <- get(paste0("adaptive_values_", plot, "_", list_name_snp_Rmarkdown[i]))
    adaptive_value_obs <- data.frame(
      ID = row.names(adaptive_value_obs),
      RDA1 = adaptive_value_obs[,1],
      RDA2 = adaptive_value_obs[,2]
    )
    
    adaptive_value_obs$Group <- "Adult"
    if (plot == "Lacanau") adaptive_value_obs$Group[grepl("^FR_PP", adaptive_value_obs$ID)] <- "Juvenile"
    if (plot == "Tocchi") adaptive_value_obs$Group[grepl("^IT_PP", adaptive_value_obs$ID)] <- "Juvenile"
    adaptive_value_obs$Pop_Age <- paste0(plot, "_", adaptive_value_obs$Group)
    
    assign(paste0("adaptive_value_obs_",plot,"_",list_name_snp_Rmarkdown[i]), adaptive_value_obs)
  }
}

#Compute density values across all datasets
for (i in seq_along(list_rda_Rmarkdown)) {
  rda_model <- get(list_rda_Rmarkdown[i])
  name_snp  <- list_name_snp_Rmarkdown[i]
  flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"
  
  lc_scores <- scores(rda_model, display = "lc", scaling = "none", choices = 1:2)
  lc_df <- as.data.frame(lc_scores)
  lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
  lc_df$Population_ID <- rownames(lc_df)
  lc_df$Country <- meta_data_pop_order$Country
  
  adaptive_values_range_wide_final <- data.frame(
    ID= lc_df$Population_ID,
    RDA1=lc_df$RDA1,
    RDA2=lc_df$RDA2,
    Group="range_wide",
    Pop_Age=lc_df$Country
  )
  
  adaptive_obs_Lacanau <- get(paste0("adaptive_value_obs_Lacanau_",name_snp))
  adaptive_obs_Tocchi <- get(paste0("adaptive_value_obs_Tocchi_",name_snp))
  
  adaptive_value_obs_final <- rbind(adaptive_obs_Lacanau, adaptive_obs_Tocchi, adaptive_values_range_wide_final)
  
  all_density_values <- c(all_density_values,
                          density(adaptive_value_obs_final$RDA1)$y,
                          density(adaptive_value_obs_final$RDA2)$y)
}

max_density <- max(all_density_values) * 1.05

#Main RDA plotting loop
for (i in seq_along(list_rda_Rmarkdown)) {
  rda_model <- get(list_rda_Rmarkdown[i])
  name_snp <- list_name_snp_Rmarkdown[i]
  flip_RDA2 <- list_rda_Rmarkdown[i] == "RDA_all_outliers_MC"
  
  explained_variance_round <- round(rda_model$CCA$eig*100/sum(rda_model$CCA$eig), 1)
  
  #Environmental arrows
  env_scores <- scores(rda_model, display = "bp")
  env_df <- as.data.frame(env_scores)
  env_df$RDA2 <- flip_if_needed(env_df$RDA2, flip_RDA2)
  env_df$RDA1 <- env_df$RDA1 * 0.15
  env_df$RDA2 <- env_df$RDA2 * 0.15
  env_df$Variable <- rownames(env_df)
  
  adaptive_obs_Lacanau <- get(paste0("adaptive_value_obs_Lacanau_",name_snp))
  adaptive_obs_Tocchi <- get(paste0("adaptive_value_obs_Tocchi_",name_snp))
  adaptive_obs_Lacanau$RDA2 <- flip_if_needed(adaptive_obs_Lacanau$RDA2, flip_RDA2)
  adaptive_obs_Tocchi$RDA2 <- flip_if_needed(adaptive_obs_Tocchi$RDA2, flip_RDA2)
  
  lc_scores <- scores(rda_model, display = "lc", scaling = "none", choices = 1:2)
  lc_df <- as.data.frame(lc_scores)
  lc_df$RDA2 <- flip_if_needed(lc_df$RDA2, flip_RDA2)
  lc_df$Population_ID <- rownames(lc_df)
  lc_df$Country <- meta_data_pop_order$Country
  adaptive_values_range_wide_final <- data.frame(
    ID= lc_df$Population_ID,
    RDA1=lc_df$RDA1,
    RDA2=lc_df$RDA2,
    Group="range_wide",
    Pop_Age=lc_df$Country
  )
  
  adaptive_value_obs_final <- rbind(adaptive_obs_Lacanau, adaptive_obs_Tocchi, adaptive_values_range_wide_final)
  
  rda_plot <- ggplot(adaptive_value_obs_final, aes(x = RDA1, y = RDA2, color = Pop_Age)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_segment(data = env_df, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
                 arrow = arrow(length = unit(0.3, "cm")), color = "black", size = 0.2) +
    geom_text(data = env_df, aes(x = RDA1*0.8, y = RDA2*0.8, label = Variable),
              color = "black", size = 3) +
    scale_color_manual(values = group_colors) +
    theme_minimal() +
    labs(title = paste("RDA space:", name_snp, "SNPs"),
         x = paste0("RDA1 (", explained_variance_round[1], "%)"),
         y = paste0("RDA2 (", explained_variance_round[2], "%)")) +
    theme(legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
  print(rda_plot)
}
```