---
title: "SPAGeDi_format_Ppinaster_study"
author: "Thomas Francisco"
date: "2026-01-14"
output:
  html_document:
    number_sections: true 
    toc: true 
    toc_float: true 
    collapsed: True 
    toc_depth: 3
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/tfrancisco/Documents/Thesis/")
rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
```

```{r library, echo=TRUE, message=FALSE, warning=FALSE}
library(polysat)
library(dplyr)
library(gstat)
library(sp)
library(rgdal)
```

In this script, we used SPAGeDi software to estimate inbreeding within plots and SGS analysis to study local structure and gain insight into local effective population size. We also examined juvenile assignments to identify the parents of the juveniles, particularly focusing on the few juveniles in Tocchi that seem pre-adapted to the future climate. 
We only computed Nb using adult individuals.

In this script, we transformed the genomic data to match the SPAGeDi format and generated 10 random sets of 500 SNPs used to build the models.  

# Format

## Load genomic data Lacanau and Tocchi individuals

```{r load data}
load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Lacanau_imputed.Rdata")

load("Data/Chapter3/Pinus_pinaster/GEA_model_pinaster/Genomic_data/good_genotyped_comparison_LAC_TOC/df_Tocchi_imputed.Rdata")

#Coordinates
coordinates_Lacanau <- read.csv("Data/Chapter3/Pinus_pinaster/Coordinates_individuals/Lacanau_coordinates_all_indiv.csv",h=T,sep=";",dec=",")

colnames(coordinates_Lacanau) <- c("VCF_ID","Latitude","Longitude","Cohort")

coordinates_Tocchi <- read.csv("Data/Chapter3/Pinus_pinaster/Coordinates_individuals/Tocchi_coordinates_all_indiv.csv",h=T,sep=";",dec=",")

colnames(coordinates_Tocchi) <- c("VCF_ID","Latitude","Longitude","Cohort")
```

We need to match the IDs of the coordinates and the genomic data, and the IDs should be shorter than 15 characters. We are going to match the IDs of the coordinates.  
```{r match coordinates Lacanau}
#Changed name in coordinate first
coordinates_Lacanau$VCF_ID <- sub(
  "^FRA00051_PP_SA",
  "FRA00051_s",
  coordinates_Lacanau$VCF_ID
)

coordinates_Lacanau$VCF_ID <- sub(
  "^FRA00051_PP_",
  "FRA00051_",
  coordinates_Lacanau$VCF_ID
)

#Now we can apply this to the genomic data
df_Lacanau_imputed_ok <- df_Lacanau_imputed #New file where we will stock the data with good format

rn <- rownames(df_Lacanau_imputed)

#Identify the rows to rename
idx <- grepl("^FR_PP_62_", rn)

#Extract trailing number
num <- as.integer(sub(".*_(\\d+)$", "\\1", rn[idx]))

#Build new names
new_names <- paste0(
  "FRA00051_s",
  sprintf("%03d", num - 500)
)

#Replace row names
rn[idx] <- new_names
rownames(df_Lacanau_imputed_ok) <- rn
```

```{r match coordinates Tocchi}
df_Tocchi_imputed_ok <- df_Tocchi_imputed

rn <- rownames(df_Tocchi_imputed)

idx <- grepl("^IT_PP_63", rn)

num <- as.integer(sub(".*_(\\d+)$", "\\1", rn[idx]))

new_names <- paste0(
  "ITA00019_s",
  sprintf("%03d", num - 500)
)

rn[idx] <- new_names
rownames(df_Tocchi_imputed_ok) <- rn
```

Merged coordinates:
```{r merge all coordinates}
coordinates_all <- rbind(coordinates_Lacanau,
  coordinates_Tocchi)
```

## Combined the two genomic files and formats

```{r combined genomic files}
df_Lacanau_imputed_ok$pop <- 1
df_Tocchi_imputed_ok$pop <- 2

df_all_genomic_data <- rbind(df_Lacanau_imputed_ok, df_Tocchi_imputed_ok)
```

We need to change the format 0,1,2 to two digits: 00 for NA, 11 for 0, 12 for 1 and 22 for 2
```{r format}
df_all_genomic_data_format <- df_all_genomic_data %>%
  mutate(
    across(
      -pop,
      ~ case_when(
        . == 0 ~ "11",
        . == 1 ~ "12",
        . == 2 ~ "22",
        TRUE   ~ "00"
      )
    )
  )
```

## Add coordinates info 

Add longitude and latitude information
```{r check coordinates}
test_coord <- data.frame(VCF_ID=row.names(df_all_genomic_data_format),cl2=df_all_genomic_data_format$Affx979044112)
coordinates_all_reduce <- merge(coordinates_all,test_coord,"VCF_ID")
all(rownames(df_all_genomic_data_format) %in% coordinates_all_reduce$VCF_ID)
```

```{r add coordinates}
coordinates_all_reduce$Latitude  <- as.numeric(coordinates_all_reduce$Latitude)
coordinates_all_reduce$Longitude <- as.numeric(coordinates_all_reduce$Longitude)

coord_match <- data.frame(
  VCF_ID = rownames(df_all_genomic_data_format),
  stringsAsFactors = FALSE
)

coordinates_all_reduce_final <- merge(
  coord_match,
  coordinates_all_reduce,
  by = "VCF_ID",
  all.x = TRUE
)

#Check order
all(rownames(df_all_genomic_data_format) == coordinates_all_reduce_final$VCF_ID)
```

## Build SPAGeDi file

```{r file SPAGeDi}
#INFO FIRST ROW
Nind   <- nrow(df_all_genomic_data_format)
Ncat   <- 2 #Two populations
Ncoord <- -2#Latitude + longitude following manual
Nloci  <- ncol(df_all_genomic_data_format) - 1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- "0" #No distance intervals

#SNP names
locus_names <- colnames(df_all_genomic_data_format)[
  colnames(df_all_genomic_data_format) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#File body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_all_genomic_data_format),
  Pop   = df_all_genomic_data_format$pop,
  Lat   = coordinates_all_reduce_final$Latitude,
  Long  = coordinates_all_reduce_final$Longitude,
  df_all_genomic_data_format[ , !names(df_all_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)
```

Save file
```{r write the file}
###Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_LAC_TOC.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

# One file per population and separating the cohorts

```{r format one file per pop}
df_Lacanau_genomic_data_format <- df_Lacanau_imputed_ok %>%
  mutate(
    across(
      -pop,
      ~ case_when(
        . == 0 ~ "11",
        . == 1 ~ "12",
        . == 2 ~ "22",
        TRUE   ~ "00"
      )
    )
  )

#Remove pop id that is not usefull here: 
df_Lacanau_genomic_data_format <- df_Lacanau_genomic_data_format[,-8470]

df_Tocchi_genomic_data_format <- df_Tocchi_imputed_ok %>%
  mutate(
    across(
      -pop,
      ~ case_when(
        . == 0 ~ "11",
        . == 1 ~ "12",
        . == 2 ~ "22",
        TRUE   ~ "00"
      )
    )
  )

df_Tocchi_genomic_data_format <- df_Tocchi_genomic_data_format[,-8470]
```

## Add coordinates info 

Add longitude and latitude information
```{r check coordinates one file per pop}
#Lacanau
test_coord_LAC <- data.frame(VCF_ID=row.names(df_Lacanau_genomic_data_format),cl2=df_Lacanau_genomic_data_format$Affx979044112)
coordinates_all_reduce_LAC <- merge(coordinates_Lacanau,test_coord,"VCF_ID")
all(rownames(df_Lacanau_genomic_data_format) %in% coordinates_all_reduce_LAC$VCF_ID)
#Remove extra column
coordinates_all_reduce_LAC <- coordinates_all_reduce_LAC[,-5]

#Tocchi
test_coord_TOC <- data.frame(VCF_ID=row.names(df_Tocchi_genomic_data_format),cl2=df_Tocchi_genomic_data_format$Affx979044112)
coordinates_all_reduce_TOC <- merge(coordinates_Tocchi,test_coord,"VCF_ID")
all(rownames(df_Tocchi_genomic_data_format) %in% coordinates_all_reduce_TOC$VCF_ID)
#Remove extra column
coordinates_all_reduce_TOC <- coordinates_all_reduce_TOC[,-5]
```

```{r add coordinates one file per pop}
list_plots <- c("Lacanau","Tocchi")
list_abbrev <- c("LAC","TOC")

for(x in 1:length(list_plots)){
  
  plot <- list_plots[x]
  name_abbrev <- list_abbrev[x]
  
  coordinate_df <- get(paste0("coordinates_all_reduce_",name_abbrev))
  df_genomic_data <- get(paste0("df_",plot,"_genomic_data_format"))
  
coordinate_df$Latitude  <- as.numeric(coordinate_df$Latitude)
coordinate_df$Longitude <- as.numeric(coordinate_df$Longitude)

coord_match <- data.frame(
  VCF_ID = rownames(df_genomic_data),
  stringsAsFactors = FALSE
)

coordinates_all_reduce_final <- merge(
  coord_match,
  coordinate_df,
  by = "VCF_ID",
  all.x = TRUE
)

#Check order
all(rownames(df_genomic_data) == coordinates_all_reduce_final$VCF_ID)

colnames(coordinates_all_reduce_final) <- c("VCF_ID","Latitude","Longitude","pop")

assign(paste0("coordinates_",plot,"_reduce_final"),coordinates_all_reduce_final)
  
}

#Add the pop info (cohort) in the genomic file
df_Lacanau_genomic_data_format$pop <- coordinates_Lacanau_reduce_final$pop

df_Tocchi_genomic_data_format$pop <- coordinates_Tocchi_reduce_final$pop
```

## Build SPAGeDi file

LACANAU
```{r file SPAGeDi LACANAU one file per pop}
coordinates_Lacanau_reduce_final
#LACANAU

# INFO FIRST ROW
Nind   <- nrow(df_Lacanau_genomic_data_format)
Ncat   <- 2 #Two populations
Ncoord <- -2      #Latitude + longitude following manual
Nloci  <- ncol(df_Lacanau_genomic_data_format) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- "0" #No distance intervals

 #SNP names
locus_names <- colnames(df_Lacanau_genomic_data_format)[
  colnames(df_Lacanau_genomic_data_format) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#File body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_Lacanau_genomic_data_format),
  Pop   = df_Lacanau_genomic_data_format$pop,
  Lat   = coordinates_Lacanau_reduce_final$Latitude,
  Long  = coordinates_Lacanau_reduce_final$Longitude,
  df_Lacanau_genomic_data_format[ , !names(df_Lacanau_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)

##Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_LAC_cohorts.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

Tocchi
```{r file SPAGeDi Tocchi one file per pop}
#INFO FIRST ROW
Nind   <- nrow(df_Tocchi_genomic_data_format)
Ncat   <- 2 #Two populations
Ncoord <- -2      #Latitude + longitude folling manual
Nloci  <- ncol(df_Tocchi_genomic_data_format) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- "0" #No distance intervals

 #SNP names
locus_names <- colnames(df_Tocchi_genomic_data_format)[
  colnames(df_Tocchi_genomic_data_format) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#file body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_Tocchi_genomic_data_format),
  Pop   = df_Tocchi_genomic_data_format$pop,
  Lat   = coordinates_Tocchi_reduce_final$Latitude,
  Long  = coordinates_Tocchi_reduce_final$Longitude,
  df_Tocchi_genomic_data_format[ , !names(df_Tocchi_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)

##Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_TOC_cohorts.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

# Combining all pop in one file and separate cohorts

Add the cohorts info in the genomic file instead of population information:
```{r combined in one df all cohorts and pop}
df_all_genomic_data_format$pop <- paste0(coordinates_all_reduce_final$Cohort,"_",df_all_genomic_data_format$pop)

coordinates_transformation <- coordinates_all_reduce_final
colnames(coordinates_transformation) <- c("VCF_ID","y","x","Cohort")

coordinates(coordinates_transformation) <- ~ y + x

coord_ok <- coordinates_transformation %>% 
  sf::st_as_sf( coords = c("y", "x") ) %>%
  sf::st_set_crs( 4326 ) %>%      #Current CRS is WSG84
  sf::st_transform( 32632 ) %>%   #Transform CRS to 32632
  sf::st_coordinates()  #Export new coordinates

coordinates_transformation_UTM <- coord_ok

coordinates_transformation_UTM_final <-data.frame(cbind(coordinates_all_reduce_final[,c(1)],coordinates_transformation_UTM,coordinates_all_reduce_final[,c(4)]))
colnames(coordinates_transformation_UTM_final) <- c("VCF_ID","Latitude","Longitude","Cohort")
```

## Build SPAGeDi file

```{r file SPAGeDi tot, eval=FALSE, include=FALSE}
#INFO FIRST ROW
Nind   <- nrow(df_all_genomic_data_format)
Ncat   <- 4 #Two populations
Ncoord <- 2      #Latitude + longitude folling manual
Nloci  <- ncol(df_all_genomic_data_format) - 1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- "-9" #No distance intervals

 #SNP names
locus_names <- colnames(df_all_genomic_data_format)[
  colnames(df_all_genomic_data_format) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#File body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_all_genomic_data_format),
  Pop   = df_all_genomic_data_format$pop,
  Lat   = coordinates_transformation_UTM_final$Latitude,
  Long  = coordinates_transformation_UTM_final$Longitude,
  df_all_genomic_data_format[ , !names(df_all_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)
```

Save file
```{r write the file tot, eval=FALSE, include=FALSE}
##Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_LAC_TOC_cohorts_sep_4pop_UTM_coord.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

# Test file with few individual and SNP

The number of SNP is too big so we need to reduce it, we can randomly select 500 snps to run the analyses and do it 10 times to account for uncertanty in the SNP selection
```{r file SPAGeDi tot reduce set, eval=FALSE, include=FALSE}
set.seed(123)

n_random <- 500

total_cols <- ncol(df_all_genomic_data_format)

#Randomly sample from all columns EXCEPT the last one
random_cols <- sample(seq_len(total_cols - 1), n_random)

#Add the last column
selected_cols <- c(random_cols, total_cols)

#Subset dataframe
df_all_genomic_data_format_2000 <- df_all_genomic_data_format[, selected_cols]

#INFO FIRST ROW
Nind   <- nrow(df_all_genomic_data_format_2000)
Ncat   <- 4 #Two populations
Ncoord <- 2      #Latitude + longitude folling manual
Nloci  <- ncol(df_all_genomic_data_format_2000) - 1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- "-9" #No distance intervals

 #SNP names
locus_names <- colnames(df_all_genomic_data_format_2000)[
  colnames(df_all_genomic_data_format_2000) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#File body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_all_genomic_data_format_2000),
  Pop   = df_all_genomic_data_format_2000$pop,
  Lat   = coordinates_transformation_UTM_final$Latitude,
  Long  = coordinates_transformation_UTM_final$Longitude,
  df_all_genomic_data_format_2000[ , !names(df_all_genomic_data_format_2000) %in% "pop"],
  stringsAsFactors = FALSE
)
```

Save file
```{r write the file tot reduce snps, eval=FALSE, include=FALSE}
##Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_LAC_TOC_cohorts_sep_4pop_UTM_coord_2000.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

# SPAGeDi file without outliers

To estimate the spatial genetic structure (SGS), we should "assume that genetic markers are neutral and the impact of selection will not be addressed" Vekemans and Hardy 2004. Therefore, we are going to remove the SNP identified in the 'all MC outlier' set to try to fulfil this assumption.

Additionally, we are going to focus on adults for SGS because neighbourhood size appears to be more closely associated with effective dispersal among breeders than offsprings, and we will focus on each individually because they are very geographically and genetically distant. 
To summarise, We are going to produce one genomic file per population containing only adults and no climate-associated SNPs. 
SPAGeDi cannot run on my computer with more than 500 SNPs, so we are going to randomly sample 500 SNPs ten times and analyse the effect of the SNP set. We might sum up the results if they are not too different. 

```{r remove outliers SNPs}
#Load outlier SNPs
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/all_outliers_set_MC_common.Rdata")

#Lacanau
df_Lacanau_genomic_data_format_wo_outliers <- df_Lacanau_genomic_data_format[,!colnames(df_Lacanau_genomic_data_format) %in% all_outliers_set_MC_common]
#Keep only the adults
df_Lacanau_genomic_data_format_wo_outliers_adults <- df_Lacanau_genomic_data_format_wo_outliers %>% filter(pop=="Adult")

#Tocchi
df_Tocchi_genomic_data_format_wo_outliers <- df_Tocchi_genomic_data_format[,!colnames(df_Tocchi_genomic_data_format) %in% all_outliers_set_MC_common]

df_Tocchi_genomic_data_format_wo_outliers_adults <- df_Tocchi_genomic_data_format_wo_outliers %>% filter(pop=="Adult")
```

Change the coordinates in meters. 
```{r coordinates}
#Keep only the adults
coordinates_Lacanau_reduce_final_adults <- coordinates_Lacanau_reduce_final %>% filter(pop == "Adult")
coordinates_Tocchi_reduce_final_adults <- coordinates_Tocchi_reduce_final %>% filter(pop=="Adult")

#Change the coordinates
list_plots <- c("Lacanau","Tocchi")
for(x in 1:length(list_plots)){
  
  plot_name <- list_plots[x]
  
  coordinates_reduce_final_adults <- get(paste0("coordinates_",plot_name,"_reduce_final_adults"))
  
coordinates_transformation <- coordinates_reduce_final_adults
colnames(coordinates_transformation) <- c("VCF_ID","y","x","Cohort")

coordinates(coordinates_transformation) <- ~ y + x

coord_ok <- coordinates_transformation %>% 
  sf::st_as_sf( coords = c("y", "x") ) %>%
  sf::st_set_crs( 4326 ) %>%      #Current CRS is WSG84
  sf::st_transform( 32632 ) %>%   #Transform CRS to 32632
  sf::st_coordinates()  #Export new coordinates

coordinates_transformation_UTM <- coord_ok

coordinates_transformation_UTM_final <-data.frame(cbind(coordinates_reduce_final_adults[,c(1)],coordinates_transformation_UTM,coordinates_reduce_final_adults[,c(4)]))
colnames(coordinates_transformation_UTM_final) <- c("VCF_ID","Latitude","Longitude","Cohort")

  assign(paste0("coordinates_transformation_UTM_final_",plot_name),coordinates_transformation_UTM_final)
}
```

Lacanau
```{r file SPAGeDi LACANAU no outliers reduce SNPs}
##10 random df
list_seed <- seq(1,10,by=1)
list_plot <- c("Lacanau","Tocchi")
list_header_classes <- c(paste(
  c("6", "20", "50","80","100","130","160"),
  collapse = "\t"
),paste(
  c("6", "20", "50","80","100","130","160"),
  collapse = "\t"
))

for(x in 1:length(list_plot)){
  
  plot_name <- list_plot[x]
  
  genomic_df <- get(paste0("df_",plot_name,"_genomic_data_format_wo_outliers_adults"))
  coordinate_df <- get(paste0("coordinates_transformation_UTM_final_",plot_name))
  
  header_classes <- list_header_classes[x]
  for(i in 1:length(list_seed)){
    
    #Select 500 random SNPs
    set.seed(i)
    n_random <- 500
    total_cols <- ncol(genomic_df)

#Randomly sample from all columns EXCEPT the last one
random_cols <- sample(seq_len(total_cols - 1), n_random)

#Add the last column
selected_cols <- c(random_cols, total_cols)

#Subset dataframe
df_all_genomic_data_format_500 <- genomic_df[, selected_cols]

#Built the SPAGeDi file
#INFO FIRST ROW
Nind   <- nrow(df_all_genomic_data_format_500)
Ncat   <- 1 #One populations
Ncoord <- 2      #Latitude + longitude following manual
Nloci  <- ncol(df_all_genomic_data_format_500) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

##SECOND ROW
header2 <- header_classes #No distance intervals

 #SNP names
locus_names <- colnames(df_all_genomic_data_format_500)[
  colnames(df_all_genomic_data_format_500) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#File body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_all_genomic_data_format_500),
  Pop   = df_all_genomic_data_format_500$pop,
  Lat   = coordinate_df$Latitude,
  Long  = coordinate_df$Longitude,
  df_all_genomic_data_format_500[ , !names(df_all_genomic_data_format_500) %in% "pop"],
  stringsAsFactors = FALSE
)

##Write file in R using writeLines and write.table 

#Where to save it
outfile <- paste0("Data/Chapter3/Pinus_pinaster/SPAGeDi/file/Per_plot_Adults/",plot_name,"/SPAGeDi_500snps_wo_outliers_only_adults_seed",i,".txt")

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
  }
}
```

Tocchi
```{r file SPAGeDi Tocchi wo outliers}
# INFO FIRST ROW
Nind   <- nrow(df_Tocchi_genomic_data_format)
Ncat   <- 2 #Two populations
Ncoord <- -2      #Latitude + longitude folling manual
Nloci  <- ncol(df_Tocchi_genomic_data_format) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

##FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

### SECOND ROW
header2 <- "0" #No distance intervals

 #SNP names
locus_names <- colnames(df_Tocchi_genomic_data_format)[
  colnames(df_Tocchi_genomic_data_format) != "pop"
]

##THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#file body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_Tocchi_genomic_data_format),
  Pop   = df_Tocchi_genomic_data_format$pop,
  Lat   = coordinates_Tocchi_reduce_final$Latitude,
  Long  = coordinates_Tocchi_reduce_final$Longitude,
  df_Tocchi_genomic_data_format[ , !names(df_Tocchi_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)

###Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_TOC_cohorts.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```


# Pairwise kinship

We can use the SPAGeDi tool to calculate the kinship between individuals (both adults and juveniles) in Lacanau and Tocchi. This information could help us identify the parents of outlier juveniles, which would also be useful for the simulation analysis.    
```{r remove outliers SNPs kinship}
#load outlier SNPs
load("Results/Chapter3/Pinaster/GEA_model_pinaster/GEA/outliers/set/all_outliers_set_MC_common.Rdata")

#Lacanau
df_Lacanau_genomic_data_format_wo_outliers <- df_Lacanau_genomic_data_format[,!colnames(df_Lacanau_genomic_data_format) %in% all_outliers_set_MC_common]
#Keep only the adults
#df_Lacanau_genomic_data_format_wo_outliers_adults <- df_Lacanau_genomic_data_format_wo_outliers %>% filter(pop=="Adult")

#Tocchi
df_Tocchi_genomic_data_format_wo_outliers <- df_Tocchi_genomic_data_format[,!colnames(df_Tocchi_genomic_data_format) %in% all_outliers_set_MC_common]

#df_Tocchi_genomic_data_format_wo_outliers_adults <- df_Tocchi_genomic_data_format_wo_outliers %>% filter(pop=="Adult")
```

Change the coordinates in meters. 
```{r coordinates kinship}
#Change the coordinates
list_plots <- c("Lacanau","Tocchi")
for(x in 1:length(list_plots)){
  
  plot_name <- list_plots[x]
  
  coordinates_reduce_final_adults <- get(paste0("coordinates_",plot_name,"_reduce_final"))
  
coordinates_transformation <- coordinates_reduce_final_adults
colnames(coordinates_transformation) <- c("VCF_ID","y","x","Cohort")

coordinates(coordinates_transformation) <- ~ y + x

coord_ok <- coordinates_transformation %>% 
  sf::st_as_sf( coords = c("y", "x") ) %>%
  sf::st_set_crs( 4326 ) %>%      #current CRS is WSG84
  sf::st_transform( 32632 ) %>%   #transform CRS to 32632
  sf::st_coordinates()  #export new coordinates

coordinates_transformation_UTM <- coord_ok

coordinates_transformation_UTM_final <-data.frame(cbind(coordinates_reduce_final_adults[,c(1)],coordinates_transformation_UTM,coordinates_reduce_final_adults[,c(4)]))
colnames(coordinates_transformation_UTM_final) <- c("VCF_ID","Latitude","Longitude","Cohort")

  assign(paste0("coordinates_transformation_UTM_final_",plot_name),coordinates_transformation_UTM_final)
}
```

Lacanau
```{r file SPAGeDi LACANAU}
##10 random df
list_seed <- seq(1,10,by=1)
list_plot <- c("Lacanau","Tocchi")
list_header_classes <- c(paste(
  c("6", "20", "50","80","100","130","160"),
  collapse = "\t"
),paste(
  c("6", "20", "50","80","100","130","160"),
  collapse = "\t"
))

for(x in 1:length(list_plot)){
  
  plot_name <- list_plot[x]
  
  genomic_df <- get(paste0("df_",plot_name,"_genomic_data_format_wo_outliers"))
  coordinate_df <- get(paste0("coordinates_transformation_UTM_final_",plot_name))
  
  header_classes <- list_header_classes[x]
  for(i in 1:length(list_seed)){
    
    #Select 500 random SNPs
    set.seed(i)
    n_random <- 500
    total_cols <- ncol(genomic_df)

#Randomly sample from all columns EXCEPT the last one
random_cols <- sample(seq_len(total_cols - 1), n_random)

#Add the last column
selected_cols <- c(random_cols, total_cols)

#Subset dataframe
df_all_genomic_data_format_500 <- genomic_df[, selected_cols]

#Built the SPAGeDi file
# INFO FIRST ROW
Nind   <- nrow(df_all_genomic_data_format_500)
Ncat   <- 2 #One populations
Ncoord <- 2      #Latitude + longitude following manual
Nloci  <- ncol(df_all_genomic_data_format_500) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

### FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

### SECOND ROW
header2 <- header_classes #No distance intervals

 #SNP names
locus_names <- colnames(df_all_genomic_data_format_500)[
  colnames(df_all_genomic_data_format_500) != "pop"
]

### THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#file body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_all_genomic_data_format_500),
  Pop   = df_all_genomic_data_format_500$pop,
  Lat   = coordinate_df$Latitude,
  Long  = coordinate_df$Longitude,
  df_all_genomic_data_format_500[ , !names(df_all_genomic_data_format_500) %in% "pop"],
  stringsAsFactors = FALSE
)

###Write file in R using writeLines and write.table 

#Where to save it
outfile <- paste0("Data/Chapter3/Pinus_pinaster/SPAGeDi/file/Kinship/",plot_name,"/SPAGeDi_500snps_wo_outliers_kinship_seed",i,".txt")

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
  }
}
```

Tocchi
```{r file SPAGeDi Tocchi}
# INFO FIRST ROW
Nind   <- nrow(df_Tocchi_genomic_data_format)
Ncat   <- 2 #Two populations
Ncoord <- -2      #Latitude + longitude folling manual
Nloci  <- ncol(df_Tocchi_genomic_data_format) -1 #Number loci
digits <- 1 #Only one based on manual
ploidy <- 2

### FIRST ROW
header1 <- paste(
  Nind, Ncat, Ncoord, Nloci, digits, ploidy,
  sep = "\t"
)

### SECOND ROW
header2 <- "0" #No distance intervals

 #SNP names
locus_names <- colnames(df_Tocchi_genomic_data_format)[
  colnames(df_Tocchi_genomic_data_format) != "pop"
]

### THIRD ROW
header3 <- paste(
  c("Ind", "Pop", "Lat", "Long", locus_names),
  collapse = "\t"
)

#file body row 4 
spagedi_body <- data.frame(
  Ind   = rownames(df_Tocchi_genomic_data_format),
  Pop   = df_Tocchi_genomic_data_format$pop,
  Lat   = coordinates_Tocchi_reduce_final$Latitude,
  Long  = coordinates_Tocchi_reduce_final$Longitude,
  df_Tocchi_genomic_data_format[ , !names(df_Tocchi_genomic_data_format) %in% "pop"],
  stringsAsFactors = FALSE
)

###Write file in R using writeLines and write.table 

#Where to save it
outfile <- "Data/Chapter3/Pinus_pinaster/SPAGeDi/file/spagedi_snps_TOC_cohorts.txt"

#Write the three first rows
writeLines(
  c(header1, header2, header3),
  con = outfile
)

#Write fourth row with VCFI_ID, coordinates and genotype info for each individual with 1 individual per row
write.table(
  spagedi_body,
  file = outfile,
  append = TRUE,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = FALSE
)

#VERY IMPORTANT END FILE WITH "END" see manual
write("END", file = outfile, append = TRUE) 
```

# Running SPAGeDi

Following the manual recommendations, we performed the selected choices to run the models:  
-   